### 2023-05-15 doc/paper.tex
Add whitepaper

\documentclass{article}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage[mathscr]{eucal}
\usepackage{authblk}

\title{Sizing the bets in a focused portfolio}
%\title{Capital allocation for a focused portfolio based on fundamental analysis}

\author[$\dagger$]{Vuko Vuk\v{c}evi\'{c}}

\affil[$\dagger$]{In silico Ltd., Zagreb, Croatia, vuko.vukcevic@insilico.hr}
%\affil[$\star$]{SimScale GmbH, Munich, Germany, vvukcevic@simscale.com}

\begin{document}

\maketitle

\section{Summary}
\label{sec:summary}

This paper consist of two parts, one without the mathematics and one with the
mathematics. For a non--mathematical reader, it is best to skip the two sections
related to mathematics and numerics, namely~\autoref{sec:mathematics}
and~\autoref{sec:numerics}. The reader who is interested in mathematics should
still read the preceding sections since they provide context within which the
mathematics is built on.

\section{Motivation}
\label{sec:motivation}

Before going on to explain the motivation about this work, a bit of context
might come useful. In silico is an employee--owned engineering consultancy
company that invests its excess cash into publicly traded stocks. The excess
cash comes sporadically, and In silico has been a net buyer over the last four
years: A trend that will likely continue. When the excess cash comes in, the
following question arises: How much money to put in which stocks? That question
had been answered by looking at our fundamental analysis for each company, and
performing some vague hand calculations. That of course works well because of
all the other uncertainties in the investment process, but the motivation was
to:
\begin{enumerate}
    \item Improve the odds of us behaving rationally (i.e. minimize
    psychological misjudgments such as anchoring bias~\cite{kahneman},
    consistency and commitment tendency~\cite{cialdini}, and
    others with their combined effects~\cite{almanack}),
    \item Save time (i.e. avoid doing hand calculations).
\end{enumerate}

Since mathematics is usually good at keeping people rational, and software is
great for saving time, a decision has been made to write a software that answers
the following question: For a set of candidate companies and their current
market capitalization, each having a set of scenarios defined by a probability
and intrinsic value estimate, how much of our capital to invest in each?

\section{Introduction}
\label{sec:introduction}

The answer to this question has been given by Kelly with his widely known Kelly
formula (sometimes called Kelly criterion)~\cite{kelly}. The Kelly's approach
starts by maximizing long--term growth of capital when one is presented with
infinite amount of opportunities to bet. This work extends that idea by
considering:
\begin{enumerate}
    \item Multiple companies (stocks) in parallel,
    \item An arbitrary number of scenarios for each company,
    \item A fundamentals--based analysis (after all, a stock is an ownership
    share of business~\cite{intelligentInvestor}.
\end{enumerate}

The generalization leads to a non--linear system of equations that when solved,
yields a fraction of capital to invest in each of the candidate companies. The
mathematical derivation is presented in \autoref{sec:mathematics}, while in the
next two subsections (\autoref{sec:assumptions} and \autoref{sec:safetyMargin}),
a discussion on assumptions in a non--mathematical way, and the margin of
safety, are presented.

\subsection{Disclaimers and Assumptions}
\label{sec:assumptions}

The underlying philosophy is that one should spend the majority of his time
analyzing investments and thinking about intrinsic values under different
scenarios that might play out, independently of outside thoughts and events.
However, calculating intrinsic value of a company is more of an art than a
science, especially for a high-quality, growing businesses within one's circle
of competence. And according to Charlie Munger, Warren Buffett, Mohnish Pabrai
and the like-minded others from whom the author got the inspiration for this
work, one should focus precisely on getting such great businesses for a fair
price. That means that one shouldn't take what this approach says at face value,
and one should probably use its guidance infrequently.

During the mathematical derivation presented in \autoref{sec:mathematics}, an
assumption is made that the number of bets is very high (tends to infinity).
This work does not try to justify this assumption in a strong mathematical
form~\footnote{It is on the author's TODO list to try and prove this.}. Here's a
soft, non-mathematical reasoning on why your author thinks this is fine: The
assumption is made in order to write the equations in terms of probabilities
instead of the number of outcomes divided by the number of bets. Therefore, as
long as the input probabilities are \textit{conservatively} estimated, the
framework should still be valid. This essentially represents the most important
margin of safety~\cite{intelligentInvestor}.

\subsection{Margin of Safety}
\label{sec:safetyMargin}

In addition to the most important margin of safety mentioned above, there are
however a couple of more margins of safety embedded in the current framework.
These are:
\begin{itemize}
    \item No shorting allowed. From a purely mathematical point of view,
    shorting would be allowed. Without detailed math, it's easy to see how a
    company with a negative expected return would result in a short position.
    However, due to the asymmetry of the potential losses compared to gains,
    coupled with the usual time--frame limit that comes with shorting, the
    author decided to avoid it.
    \item No use of leverage allowed. Again, from a purely mathematical point of
    view, use of leverage would sometimes be useful. The thinking in avoiding to
    use leverage is that no--one should be in a hurry to get rich, and should
    avoid risking good night's sleep based on short--term market fluctuations,
    which are fairly hard to predict consistently (unless you work at
    Renaissance Technologies, in which case you can scratch this).
    \item No company without at least one downside scenario is allowed. By
    disallowing inputs without downside, the framework forces you to focus and
    think about what can go wrong, as opposed to dreaming about what can go
    right. However, if one is absolutely (100\%) sure that a company does not
    have a downside, then the solution is to lever up as much as possible, and
    put all the money in that one company. It is our feeling that the best way
    to handle such cases is outside of this framework. Alternatively, one can
    always model an unknown downside scenario with a small probability (say
    5--20\%) and intrinsic value of zero.
\end{itemize}

These assumptions and margins of safety are embedded into the framework in order
to try and tie the rational mathematics with common--sense from other
disciplines.

\section{Mathematics}
\label{sec:mathematics}

The following derivation mostly follows the first part of the work by Byrnes and
Barnett~\cite{byrnesBarnett}. The problem statement is repeated here for
convenience: Given a set of candidate companies, each having a set of scenarios
described by the probability $p$ and estimate of the intrinsic value
$\mathcal{V}$, calculate the optimal allocation fraction $f$ for each candidate
company by maximizing the long--term growth rate of assets. After a single
outcome (realization), the change in value of assets can be written as follows:
\begin{equation}
\label{eq:1}
    \mathcal{A}_{after}
  = 
    \mathcal{A}_{before}
    \left( 1 + \sum_{j}^{N_c} f_j k_j \right)
\end{equation}

\noindent where $\mathcal{A}$ is the value of assets (capital), $N_c$ is the
number of candidate companies to consider, $f_j$ is the allocated fraction to
$j$th company, and $k_j$ is one of the returns for a company $j$ defined as the
relative difference between the estimated intrinsic value under a given scenario
($\mathcal{V}$) and the market capitalization at the time of investing
($\mathcal{M}$):
\begin{equation}
\label{eq:2}
    k_j = \frac{\mathcal{V}_j - \mathcal{M}_j}{\mathcal{M}_j}
\end{equation}

\noindent If a significant number of (re)allocations ($N_a$) is performed in
succession, the equation~\eqref{eq:1} can be written as follows:
\begin{equation}
\label{eq:3}
    \mathcal{A}_{N_a}
  = 
    \mathcal{A}_{0} \prod_{i_1, i_2, \hdots, i_{N_o}}
    \left( 1 + \sum_{j}^{N_c} f_j k_{ij} \right)^{n_i}
\end{equation}

\noindent where $n_i \in {n_{i1}, n_{i2}, \hdots, n_{N_o}} $ is the number of
the $i$th outcome that occurred $n_i$ times. Note that $k_{ij}$ represents the
return of the $j$th company in the $i$th outcome. Following original Kelly's
approach, a logarithmic growth function $\mathcal{G}$ is introduced:
\begin{equation}
\label{eq:4}
    \mathcal{G} = \lim_{N_a \to \infty} \frac{1}{N_a} \ln
        \frac{\mathcal{A}_{N_a}}{\mathcal{A}_0}
\end{equation}

\noindent and the goal is to find its maximum with respect to allocation
fractions $f_j$:
\begin{equation}
\label{eq:5}
    \frac{\partial \mathcal{G}}{\partial f_j} = 0
\end{equation}

\noindent Substituting equation~\eqref{eq:1} into equation~\eqref{eq:5} results
in the following equation, after some calculus and algebra:
\begin{equation}
\label{eq:6}
    \lim_{N_a \to \infty} \frac{1}{N_a}
    \sum_{i}^{N_o} \frac{n_i k_{ij}}{1 + \sum_{j}^{N_c} f_j k_{ij}} = 0
\end{equation}

\noindent If one assumes an infinite number of (re)allocations $N_a$,
the following relation holds\footnote{This is an assumption that your author is
uncomfortable with, but \textit{feels} it is fine because of the margins of
safety embedded into the thinking that goes into assessing each investment
opportunity.}:
\begin{equation}
\label{eq:7}
    \lim_{N_a \to \infty} \frac{n_i}{N} = p_i
\end{equation}

\noindent Where $p_i$ is the probability of the $i$th outcome. For example, if
there are two companies, each with two 50--50 scenarios, there will be four
outcomes in total with the probability of each outcome equal to 25\%. Finally,
substituting equation~\eqref{eq:7} into~\eqref{eq:6} results in a system of
equations written in terms of probabilities $p_i$, expected returns for each
company in each of the outcomes $k_{ij}$, and allocation fractions for each
company $f_j$:
\begin{equation}
\label{eq:8}
    \sum_{i}^{N_o} \frac{p_i k_{ij}}{1 + \sum_{j}^{N_c} f_j k_{ij}} = 0
\end{equation}

\noindent The equation~\eqref{eq:8} represents a non--linear system of equations
in the unknown fractions $f_j$, which when solved, should yield optimal
allocation strategy for maximizing long--term growth of capital.

\section{Numerics}
\label{sec:numerics}

\noindent The equation~\eqref{eq:8} can be written succinctly as:
\begin{equation}
\label{eq:9}
    \mathcal{F}_i(f_j) = 0
\end{equation}

\noindent Because $\mathcal{F}_i$ is a non--linear equation in $f_j$,
the Newton--Raphson method is used to find a numerical solution. The method is
iterative and starts by linearizing the equation around the previous solution
from the previous iteration:
\begin{equation}
\label{eq:10}
    \mathcal{F}_i^o + \sum_{j}^{N_c} \mathcal{J}_{ij}^o (f_j^n - f_j^o) = 0
\end{equation}

\noindent where $\mathcal{J}_{ij}$ is the Jacobian of $\mathcal{F_i}$, and
superscripts $^n$ and $^o$ denote the new value and the old value, respectively.
The Jacobian is:
\begin{equation}
\label{eq:11}
    \mathcal{J}_{ij}
  = 
    - \sum_{o}^{N_o}
      \frac{p_o k_{oi} k_{oj}}{\left(1 + \sum_{j}^{N_c} f_j k_{ij} \right)^2}
\end{equation}

\noindent where subscript for outcome $i$ has been changed to $o$ in order to be
able to write the Jacobian in the standard $ij$ notation.

The numerical procedure starts by assuming the uniform allocation across
all companies, i.e. $f_j = f = \frac{1}{N_c}$. Based on the $f_j^o$ in the
current iteration, the linear system in equation~\eqref{eq:10} is solved to find
the new solution $f_j^n$. The process is repeated until sufficient level of
accuracy is reached~\footnote{Usually two significant digits is enough
considering the uncertainty and the judgment required for the inputs.}.

A practical approach is used to prevent shorting and use of leverage. Shorting
would occur if any of the fractions is negative, and is prevented by filtering
out the candidate companies with negative expected return before solving the
equations. The use of leverage is avoided by normalizing the fractions after
obtaining a numerical solution, if the sum of all fractions exceeds 1. The
author is well--aware that this is somewhat a brute--force approach, and it will
be a topic of future work to avoid shorting and use of leverage by using
inequality constraints.

\section{Example}
\label{sec:example}

\noindent As an example, consider seven candidate companies, each with two to
four scenarios. Each scenario is represented by an intrinsic value and the
probability of the scenario happening (or intrinsic value being reached at some
point in the future). Note that how these numbers are obtained is outside of the
scope of this work, although it is important to stress that the validity and
conservative assumptions behind these numbers are probably the most important
part of an investor's job. The example inputs are presented
in~\autoref{tab:companyA} to~\autoref{tab:companyF}.

\begin{table}[!ht]
\caption{Company A with current market cap of 214B USD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 USD & 5\% \\
Bear thesis & 170B USD & 30\% \\
Base thesis & 270B USD & 50\% \\
Bull thesis & 360B USD & 15\% \\
\end{tabular}%
\label{tab:companyA}%
\end{table}%

\begin{table}[!ht]
\caption{Company B with current market cap of 306M USD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss& 0 USD & 5\% \\
Bear thesis & 300M USD & 50\% \\
Base thesis & 900M USD & 45\% \\
\end{tabular}%
\label{tab:companyB}%
\end{table}%

\begin{table}[!ht]
\caption{Company C with current market cap of 34M GBP.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 GBP & 10\% \\
Bear thesis & 33.5M GBP & 50\% \\
Base thesis & 135M GBP & 40\% \\
\end{tabular}%
\label{tab:companyC}%
\end{table}%

\begin{table}[!ht]
\caption{Company D with current market cap of 806M SGD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Bear thesis & 330M SGD & 40\% \\
Base thesis & 1B SGD & 50\% \\
Bull thesis & 4B SGD & 10\% \\
\end{tabular}%
\label{tab:companyD}%
\end{table}%

\begin{table}[!ht]
\caption{Company E with current market cap of 17.6 CAD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 CAD & 5\% \\
Bear thesis & 10M CAD & 25\% \\
Base thesis & 25M CAD & 70\% \\
\end{tabular}%
\label{tab:companyE}%
\end{table}%

\begin{table}[!ht]
\caption{Company F with current market cap of 581M USD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 USD & 20\% \\
Base thesis & 1B CAD & 80\% \\
\end{tabular}%
\label{tab:companyF}%
\end{table}%

\noindent Based on the inputs, the portfolio allocation that maximizes the
long--term growth--rate of capital is presented in~\autoref{tab:results}.

\begin{table}
\caption{Portfolio that maximizes long--term growth--rate of capital.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c|c|c|c|c}
Company & A & B & C & D & E & F \\
\hline
Allocation fractions & 13\% & 27\% & 26\% & 12\% & 12\% & 10\% \\
\end{tabular}%
\label{tab:results}%
\end{table}%

With the obtained fractions, it is easy to obtain some useful
statistics on the portfolio, namely:
\begin{itemize}
    \item Expected gain of 61 cents for every dollar invested,
    \item Cumulative probability of loss of capital of 13\%,
    \item Permanent loss of 95\% of capital with probability of 0.001\%.
\end{itemize}

\noindent The last item is particularly interesting to your author. According to
Actuarial Life Tables in~\cite{lifeTables}, the probability of your currently 33
year--old author dying within the next year is approximately 0.25\%. That is two
orders of magnitude higher than the probability of the permanent loss of capital
for this portfolio. Considering that the portfolio has six stocks, that is a
very strong argument against excessive diversification, especially if:
\begin{itemize}
    \item One thinks of stocks as ownership shares of businesses, which implies
    long--term thinking and not being bothered by market fluctuations, 
    \item One embeds a margin of safety in different scenarios for different
    companies by e.g. recognizing that both unknown and known bad things may
    happen.
\end{itemize}

\noindent The observation about excessive diversification Poor Charlie's
Almanac~\cite{almanack}.

\section{Future Work}
\label{sec:futureWork}

\noindent The future work will focus on adding a constraint for the maximum
allowable risk of the permanent loss of capital. The idea is that an investor
could input the probability of the permanent loss of capital and the fraction of
capital one is willing to risk, and the allocation will change accordingly. For
example, an investor might be fine with losing 50\% of capital with the
probability of 0.1\% as a worst--case outcome, which should be an additional
input to the system.

\clearpage

\bibliographystyle{unsrt}
\bibliography{./references}

\end{document}

### 2023-05-17 doc/paper.tex
Address paper review comments

@@
 performing some vague hand calculations. That of course works well because of
-all the other uncertainties in the investment process, but the motivation was
-to:
+all the other uncertainties in the investment process, but we were still
+motivated to:
 \begin{enumerate}
@@
 starts by maximizing long--term growth of capital when one is presented with
-infinite amount of opportunities to bet. This work extends that idea by
+an infinite amount of opportunities to bet on. This work extends that idea by
 considering:
@@
     \item A fundamentals--based analysis (after all, a stock is an ownership
-    share of business~\cite{intelligentInvestor}.
+    share of business~\cite{intelligentInvestor}).
 \end{enumerate}
@@
 form~\footnote{It is on the author's TODO list to try and prove this.}. Here's a
-soft, non-mathematical reasoning on why your author thinks this is fine: The
+soft, non-mathematical reasoning on why the author thinks this is fine: The
 assumption is made in order to write the equations in terms of probabilities
@@
 number of candidate companies to consider, $f_j$ is the allocated fraction to
-$j$th company, and $k_j$ is one of the returns for a company $j$ defined as the
-relative difference between the estimated intrinsic value under a given scenario
+$j$th company, and $k_j$ is a return for a company $j$ defined as the relative
+difference between the estimated intrinsic value under a given scenario
 ($\mathcal{V}$) and the market capitalization at the time of investing
@@
 
-\noindent where $n_i \in {n_{i1}, n_{i2}, \hdots, n_{N_o}} $ is the number of
-the $i$th outcome that occurred $n_i$ times. Note that $k_{ij}$ represents the
-return of the $j$th company in the $i$th outcome. Following original Kelly's
-approach, a logarithmic growth function $\mathcal{G}$ is introduced:
+\noindent where $n_i \in {n_{i1}, n_{i2}, \hdots, n_{N_o}}$ is the number of
+times the $i$th outcome has occurred. Note that $k_{ij}$ represents the return
+of the $j$th company for the $i$th outcome. Following original Kelly's approach,
+a logarithmic growth function $\mathcal{G}$ is introduced:
 \begin{equation}
@@
 
-\noindent If one assumes an infinite number of (re)allocations $N_a$,
-the following relation holds\footnote{This is an assumption that your author is
-uncomfortable with, but \textit{feels} it is fine because of the margins of
-safety embedded into the thinking that goes into assessing each investment
-opportunity.}:
+\noindent If one assumes an infinite number of (re)allocations
+$N_a$\footnote{The assumption regarding the infinite number of allocations is
+something that the author is slightly uncomfortable with, but \textit{feels} it
+is fine because of the margins of safety embedded into the thinking that goes
+into assessing each investment opportunity.}, the following relation holds:
 \begin{equation}
@@
 \begin{table}[!ht]
-\caption{Company E with current market cap of 17.6 CAD.}
+\caption{Company E with current market cap of 17.6M CAD.}
 \vspace{0.25cm}
@@
 Total loss & 0 USD & 20\% \\
-Base thesis & 1B CAD & 80\% \\
+Base thesis & 1B USD & 80\% \\
 \end{tabular}%
@@
 
-\noindent The last item is particularly interesting to your author. According to
-Actuarial Life Tables in~\cite{lifeTables}, the probability of your currently 33
-year--old author dying within the next year is approximately 0.25\%. That is two
-orders of magnitude higher than the probability of the permanent loss of capital
-for this portfolio. Considering that the portfolio has six stocks, that is a
-very strong argument against excessive diversification, especially if:
+\noindent The last item is particularly interesting to the author. According to
+Actuarial Life Tables in~\cite{lifeTables}, the probability of the (currently 33
+year--old) author dying within the next year is approximately 0.25\%. That is
+two orders of magnitude higher than the probability of the permanent loss of
+capital for this portfolio. Considering that the portfolio has six stocks, that
+is a very strong argument against excessive diversification, especially if:
 \begin{itemize}
@@
 
-\noindent The observation about excessive diversification Poor Charlie's
-Almanac~\cite{almanack}.
+\noindent The observation about excessive diversification is inline with the
+thoughts from the Poor Charlie's Almanack~\cite{almanack}.
 

### 2023-05-20 src/model/company.rs
Two unrelated updates:
1. Format the output nicely in the demo FE
2. Run test coverage in the pipeline since the runner should now be configured

@@ -102,7 +102,7 @@ impl Company {
         if (sum - 1.0).abs() > TOLERANCE {
             ValidationResult::PROBLEM(Problem {
                 code: "probabilities-for-all-scenarios-do-not-sum-up-to-one".to_string(),
-                message: format!("Probabilities of all scenarios do not sum up to 1. Sum = {sum}."),
+                message: format!("Probabilities of all scenarios for company {name} do not sum up to 1. Sum = {sum}.", name = self.name),
                 severity: Severity::ERROR,
             })
         } else {
@@ -323,7 +323,7 @@ mod test {
             .validate()
             .contains(&ValidationResult::PROBLEM(Problem {
                 code: "probabilities-for-all-scenarios-do-not-sum-up-to-one".to_string(),
-                message: "Probabilities of all scenarios do not sum up to 1. Sum = 0.8."
+                message: "Probabilities of all scenarios for company Some Company do not sum up to 1. Sum = 0.8."
                     .to_string(),
                 severity: Severity::ERROR,
             })));

### 2023-05-20 tests/integration.rs

@@ -220,7 +220,7 @@ fn test_allocate_with_validation_problems() {
         allocation_response.validation_problems.unwrap(),
         vec![ValidationResult::PROBLEM(Problem {
             code: "probabilities-for-all-scenarios-do-not-sum-up-to-one".to_string(),
-            message: "Probabilities of all scenarios do not sum up to 1. Sum = 0.98.".to_string(),
+            message: "Probabilities of all scenarios for company A do not sum up to 1. Sum = 0.98.".to_string(),
             severity: Severity::ERROR,
         })],
     );

### 2023-05-20 tests/integration.rs
Format

@@ -220,7 +220,8 @@ fn test_allocate_with_validation_problems() {
         allocation_response.validation_problems.unwrap(),
         vec![ValidationResult::PROBLEM(Problem {
             code: "probabilities-for-all-scenarios-do-not-sum-up-to-one".to_string(),
-            message: "Probabilities of all scenarios for company A do not sum up to 1. Sum = 0.98.".to_string(),
+            message: "Probabilities of all scenarios for company A do not sum up to 1. Sum = 0.98."
+                .to_string(),
             severity: Severity::ERROR,
         })],
     );

### 2023-07-12 src/allocation.rs
Adding support for inequality constraints and specifically maximum allowable permanent loss of capital

@@ -1,500 +1,0 @@
-use crate::analysis::{all_outcomes, Outcome};
-use crate::model::company::Company;
-use crate::model::errors::Error;
-use crate::model::portfolio::{Portfolio, PortfolioCompany};
-use nalgebra::{DMatrix, DVector};
-use num_traits::pow::Pow;
-use slog::{info, Logger};
-
-/// Tolerance for converging the fraction during Newton-Raphson iteration. Corresponds to 1%, which
-/// is more than enough given that the real uncertainty lies in the input data and not here.
-pub const FRACTION_TOLERANCE: f64 = 1e-2;
-
-/// Relaxation factor used when updating solution vector in an iteration of the nonlinear loop
-const RELAXATION_FACTOR: f64 = 0.7;
-
-/// Maximum number of iterations for the nonlinear solver
-pub const MAX_ITER: u32 = 1000;
-
-/// Calculates allocation factors (fractions) for each company based on the Kelly criterion, by
-/// solving N nonlinear equations (N = number of candidates) using the Newton-Raphson algorithm
-pub fn kelly_allocate(
-    candidates: Vec<Company>,
-    max_iter: u32,
-    logger: &Logger,
-) -> Result<Portfolio, Error> {
-    let n_companies: usize = candidates.len();
-    info!(
-        logger,
-        "Solving the Kelly allocation equations for {n_companies} companies"
-    );
-
-    // Initial guess for fractions assumes uniform allocation across all companies
-    let uniform_fraction: f64 = 1.0 / n_companies as f64;
-    let mut fractions: DVector<f64> = DVector::from_element(n_companies, uniform_fraction);
-
-    // Get all outcomes for a list of candidates. Note that the fractions are not relevant here
-    // since we only care about non-weighted company returns and probability
-    let mut portfolio: Portfolio = Portfolio {
-        companies: candidates
-            .into_iter()
-            .map(|c| PortfolioCompany {
-                company: c,
-                fraction: uniform_fraction,
-            })
-            .collect(),
-    };
-    let outcomes: Vec<Outcome> = match all_outcomes(&portfolio) {
-        Ok(o) => o,
-        Err(e) => return Err(e),
-    };
-
-    info!(logger, "Starting Newton-Raphson loop.");
-    let mut counter: u32 = 0;
-    loop {
-        // Update the fractions in the portfolio for calculating Kelly function and Jacobian
-        portfolio
-            .companies
-            .iter_mut()
[…]

### 2023-07-12 src/constraints.rs

@@ -1,0 +1,2 @@
+pub mod capital_loss_constraint;
+pub mod constraint;

### 2023-07-12 src/constraints/capital_loss_constraint.rs

@@ -1,0 +1,53 @@
+use crate::constraints::constraint::{Constraint, InequalityConstraint};
+use crate::model::portfolio::Portfolio;
+use nalgebra::DVector;
+use ordered_float::OrderedFloat;
+
+/// [CapitalLossConstraint] implements the [InequalityConstraint] interface that provides matrix
+/// contributions to the Kelly allocation problem, both for inactive and active variants.
+#[derive(Debug)]
+pub struct CapitalLossConstraint {
+    pub fraction_of_capital: f64,
+    pub probability_of_loss: f64,
+}
+
+impl InequalityConstraint for CapitalLossConstraint {}
+
+impl Constraint for CapitalLossConstraint {
+    fn d_constraint_d_fractions(&self, portfolio: &Portfolio) -> DVector<f64> {
+        DVector::from_vec(
+            portfolio
+                .companies
+                .iter()
+                .map(|p| {
+                    p.company
+                        .scenarios
+                        .iter()
+                        .map(|s| OrderedFloat(s.probability_weighted_return(p.company.market_cap)))
+                        .min()
+                        .unwrap_or_else(|| {
+                            panic!(
+                                "Did not manage to find worst case scenario for company {:?}",
+                                p.company.ticker
+                            )
+                        })
+                        .into_inner()
+                })
+                .collect(),
+        )
+    }
+
+    fn function_value(&self, portfolio: &Portfolio, slack_variable: f64) -> f64 {
+        self.d_constraint_d_fractions(portfolio)
+            .iter()
+            .enumerate()
+            .map(|(c_i, dc_df)| dc_df * portfolio.companies[c_i].fraction)
+            .sum::<f64>()
+            - self.probability_of_loss * self.fraction_of_capital
+            + slack_variable
+    }
+
+    fn is_satisfied(&self, portfolio: &Portfolio) -> bool {
+        self.function_value(portfolio, 0.0) <= 0.0
+    }
+}

### 2023-07-12 src/constraints/constraint.rs

@@ -1,0 +1,21 @@
+use crate::model::portfolio::Portfolio;
+use nalgebra::DVector;
+
+/// [InequalityConstraint] is used to tag inequality constraints.
+pub trait InequalityConstraint: Constraint {}
+
+/// TODO: Declare EqualityConstraint when the time comes
+
+/// [Constraint] is a super-trait providing the interface for calculating matrix contributions when
+/// solving the Kelly allocation problem. The only thing needed for implementing a constraint is to
+/// provide the partial derivative of the constraint function with respect to allocation fractions.
+pub trait Constraint {
+    /// Partial derivative of the constraint with respect to the fractions. Ends up in the matrix.
+    fn d_constraint_d_fractions(&self, portfolio: &Portfolio) -> DVector<f64>;
+
+    /// Constraint function value. Ends up in the right-hand-side of the system.
+    fn function_value(&self, portfolio: &Portfolio, slack_variable: f64) -> f64;
+
+    /// Whether this constraint is satisfied. Used to check whether we found a viable solution.
+    fn is_satisfied(&self, portfolio: &Portfolio) -> bool;
+}

### 2023-07-12 src/model.rs

@@ -1,3 +1,4 @@
+pub mod capital_loss;
 pub mod company;
 pub mod errors;
 pub mod portfolio;

### 2023-07-12 src/model/portfolio.rs

@@ -1,3 +1,4 @@
+use crate::model::capital_loss::CapitalLoss;
 use crate::model::company::Company;
 use schemars::JsonSchema;
 use serde::{Deserialize, Serialize};
@@ -15,8 +16,9 @@ pub struct PortfolioCompany {
     pub fraction: f64,
 }
 
-/// Portfolio candidates has a list of companies
+/// Allocation input consists of a list of candidate companies and additional settings
 #[derive(Serialize, Deserialize, JsonSchema, Clone, Debug)]
-pub struct PortfolioCandidates {
-    pub companies: Vec<Company>,
+pub struct AllocationInput {
+    pub candidates: Vec<Company>,
+    pub max_permanent_loss_of_capital: Option<CapitalLoss>,
 }

### 2023-07-12 src/model/responses.rs

@@ -4,7 +4,7 @@ use crate::validation::result::ValidationResult;
 use schemars::JsonSchema;
 use serde::{Deserialize, Serialize};
 
-/// Response of the call to the allocate endpoint, contains results of both allocation and analysis
+/// Response of the call to the allocate endpoint, contains results of both allocation and analysis.
 #[derive(Serialize, Deserialize, JsonSchema, Clone, Debug)]
 pub struct AllocationResponse {
     pub result: Option<AllocationResult>,
@@ -12,38 +12,39 @@ pub struct AllocationResponse {
     pub error: Option<Error>,
 }
 
-/// Response of the call to the analyze endpoint
+/// Response of the call to the analyze endpoint.
 #[derive(Serialize, Deserialize, JsonSchema, Clone, Debug)]
 pub struct AnalysisResponse {
     pub result: Option<AnalysisResult>,
     pub error: Option<Error>,
 }
 
-/// Allocation result includes tickers and their fractions
+/// Allocation result includes tickers and their fractions.
 #[derive(Serialize, Deserialize, JsonSchema, Clone, Debug)]
 pub struct AllocationResult {
     pub allocations: Vec<TickerAndFraction>,
     pub analysis: AnalysisResult,
 }
 
-/// Analysis result includes some statistics for a given portfolio
+/// Analysis result includes some statistics for a given portfolio.
 #[derive(Serialize, Deserialize, JsonSchema, Clone, Debug)]
 pub struct AnalysisResult {
-    pub worst_case_outcome: ProbabilityAndReturn,
+    pub worst_case_outcome: ProbabilityAndReturns,
     pub cumulative_probability_of_loss: f64,
     pub expected_return: f64,
 }
 
-/// A ticker and a fraction used for minimalistic representation of the allocation calculation
+/// A ticker and a fraction used for minimalistic representation of the allocation calculation.
 #[derive(Serialize, Deserialize, JsonSchema, Clone, Debug)]
 pub struct TickerAndFraction {
     pub ticker: Ticker,
     pub fraction: f64,
 }
 
-/// Probability and return used to minimally represent an outcome
+/// Probability and returns used to minimally represent an outcome.
 #[derive(Serialize, Deserialize, JsonSchema, Clone, Debug)]
-pub struct ProbabilityAndReturn {
+pub struct ProbabilityAndReturns {
     pub probability: f64,
-    pub weighted_return: f64,
+    pub portfolio_return: f64,
+    pub probability_weighted_return: f64,
 }

### 2023-07-12 tests/integration.rs

@@ -1,7 +1,7 @@
-use charlie::allocation::{kelly_allocate, FRACTION_TOLERANCE, MAX_ITER};
 use charlie::env::create_test_logger;
+use charlie::kelly_allocation::{KellyAllocator, MAX_ITER, SOLVER_TOLERANCE};
 use charlie::model::errors::Error;
-use charlie::model::portfolio::PortfolioCandidates;
+use charlie::model::portfolio::AllocationInput;
 use charlie::model::responses::{AllocationResponse, AnalysisResponse, TickerAndFraction};
 use charlie::utils::assert_close;
 use charlie::validation::result::{Problem, Severity, ValidationResult};
@@ -9,10 +9,10 @@ use charlie::{allocate, analyze, validate};
 use slog::info;
 
 /// Make assertion tolerance the same as the fraction tolerance (no point in more accuracy)
-const ASSERTION_TOLERANCE: f64 = FRACTION_TOLERANCE;
+const ASSERTION_TOLERANCE: f64 = SOLVER_TOLERANCE;
 
 const TEST_YAML: &str = "
-          companies:
+          candidates:
           - name: A
             ticker: A
             description: Business A
@@ -119,99 +119,90 @@ const TEST_YAML: &str = "
 
 #[test]
 fn test_create_candidates_and_validate() {
-    let candidates: PortfolioCandidates = serde_yaml::from_str(&TEST_YAML.to_string()).unwrap();
+    let input: AllocationInput = serde_yaml::from_str(&TEST_YAML.to_string()).unwrap();
 
-    assert_eq!(candidates.companies.len(), 6);
+    assert_eq!(input.candidates.len(), 6);
 
     // First company
-    assert_eq!(candidates.companies[0].name, "A");
-    assert_eq!(candidates.companies[0].ticker, "A");
-    assert_eq!(candidates.companies[0].description, "Business A");
-    assert_eq!(candidates.companies[0].market_cap, 238.0e9);
+    assert_eq!(input.candidates[0].name, "A");
+    assert_eq!(input.candidates[0].ticker, "A");
+    assert_eq!(input.candidates[0].description, "Business A");
+    assert_eq!(input.candidates[0].market_cap, 238.0e9);
 
     // Scenarios for the first company
-    assert_eq!(candidates.companies[0].scenarios.len(), 4);
+    assert_eq!(input.candidates[0].scenarios.len(), 4);
     assert_eq!(
[…]

### 2023-07-12 src/constraints.rs
Constraints for no shorting

@@ -1,2 +1,3 @@
 pub mod capital_loss_constraint;
 pub mod constraint;
+pub mod no_shorting_constraint;

### 2023-07-12 src/constraints/capital_loss_constraint.rs

@@ -3,8 +3,10 @@ use crate::model::portfolio::Portfolio;
 use nalgebra::DVector;
 use ordered_float::OrderedFloat;
 
-/// [CapitalLossConstraint] implements the [InequalityConstraint] interface that provides matrix
-/// contributions to the Kelly allocation problem, both for inactive and active variants.
+/// [CapitalLossConstraint] that puts an upper bound on the permanent loss of capital the investor
+/// is comfortable with. It essentially limits the fractions such that the probability-weighted
+/// worst-case scenario doesn't exceed the specified value.
+/// TODO: Fields can be represented into a single number.
 #[derive(Debug)]
 pub struct CapitalLossConstraint {
     pub fraction_of_capital: f64,
@@ -45,9 +47,5 @@ impl Constraint for CapitalLossConstraint {
             .sum::<f64>()
             - self.probability_of_loss * self.fraction_of_capital
             + slack_variable
-    }
-
-    fn is_satisfied(&self, portfolio: &Portfolio) -> bool {
-        self.function_value(portfolio, 0.0) <= 0.0
     }
 }

### 2023-07-12 src/constraints/constraint.rs

@@ -1,10 +1,17 @@
 use crate::model::portfolio::Portfolio;
 use nalgebra::DVector;
 
-/// [InequalityConstraint] is used to tag inequality constraints.
-pub trait InequalityConstraint: Constraint {}
+/// [InequalityConstraint] extends the [Constraint] interface, adding only a default function for
+/// checking whether the inequality constraint is satisfied
+pub trait InequalityConstraint: Constraint {
+    /// Whether this constraint is satisfied. Used to check whether we found a viable solution.
+    /// All inequality constraints have to be cast into this (not-greater than 0) form.
+    fn is_satisfied(&self, portfolio: &Portfolio) -> bool {
+        self.function_value(portfolio, 0.0) <= 0.0
+    }
+}
 
-/// TODO: Declare EqualityConstraint when the time comes
+/// TODO: Declare EqualityConstraint here when the time comes.
 
 /// [Constraint] is a super-trait providing the interface for calculating matrix contributions when
 /// solving the Kelly allocation problem. The only thing needed for implementing a constraint is to
@@ -15,7 +22,4 @@ pub trait Constraint {
 
     /// Constraint function value. Ends up in the right-hand-side of the system.
     fn function_value(&self, portfolio: &Portfolio, slack_variable: f64) -> f64;
-
-    /// Whether this constraint is satisfied. Used to check whether we found a viable solution.
-    fn is_satisfied(&self, portfolio: &Portfolio) -> bool;
 }

### 2023-07-12 src/constraints/no_shorting_constraint.rs

@@ -1,0 +1,25 @@
+use crate::constraints::constraint::{Constraint, InequalityConstraint};
+use crate::model::portfolio::Portfolio;
+use nalgebra::DVector;
+
+/// [NoShortingConstraint] puts a constraint that disallows shorting. It essentially limits the
+/// fractions such that they are always non-negative.
+#[derive(Debug)]
+pub struct NoShortingConstraint {
+    /// Index representing the fraction it constraints
+    pub fraction_index: usize,
+}
+
+impl InequalityConstraint for NoShortingConstraint {}
+
+impl Constraint for NoShortingConstraint {
+    fn d_constraint_d_fractions(&self, portfolio: &Portfolio) -> DVector<f64> {
+        let mut derivative: DVector<f64> = DVector::zeros(portfolio.companies.len());
+        derivative[self.fraction_index] = -1.0;
+        derivative
+    }
+
+    fn function_value(&self, portfolio: &Portfolio, slack_variable: f64) -> f64 {
+        -portfolio.companies[self.fraction_index].fraction + slack_variable
+    }
+}

### 2023-07-12 src/kelly_allocation.rs

@@ -12,6 +12,7 @@ use slog::{info, Logger};
 use crate::analysis::{all_outcomes, worst_case_outcome, Outcome};
 use crate::constraints::capital_loss_constraint::CapitalLossConstraint;
 use crate::constraints::constraint::InequalityConstraint;
+use crate::constraints::no_shorting_constraint::NoShortingConstraint;
 use crate::model::capital_loss::CapitalLoss;
 use crate::model::company::{Company, TOLERANCE};
 use crate::model::errors::Error;
@@ -47,6 +48,48 @@ impl<'a> KellyAllocator<'a> {
         }
     }
 
+    /// Return a new [KellyAllocator] with a constraint for no shorting, for all company candidates.
+    /// The contents of the original object are moved into the new one. Panics in case a single
+    /// constraint of this type is already present.
+    pub fn with_no_shorting_constraint(self, n_candidates: usize) -> KellyAllocator<'a> {
+        info!(
+            self.logger,
+            "Setting no shorting constraints for all of the {n_candidates}."
+        );
+
+        if n_candidates < 1 {
+            panic!("Can't make sense of no shorting constraints because there are {n_candidates} candidates.")
+        }
+
+        // Fractions are always the first set of unknowns in the system.
+        let no_shorting_constraints: Vec<Box<dyn InequalityConstraint>> = (0..n_candidates)
+            .map(|i| {
+                Box::new(NoShortingConstraint { fraction_index: i })
+                    as Box<dyn InequalityConstraint>
+            })
+            .collect();
+
+        if self
+            .inequality_constraints
+            .iter()
+            .any(|c| c.type_id() == no_shorting_constraints[0].type_id())
+        {
+            panic!(
+                "Kelly allocator already initialized with no shorting constraints. Did you call \
+                with_no_shorting_constraint twice?"
+            )
+        }
+
+        let mut new_constraints = self.inequality_constraints;
+        new_constraints.extend(no_shorting_constraints);
+
+        KellyAllocator {
+            logger: self.logger,
+            max_iter: self.max_iter,
+            inequality_constraints: new_constraints,
+        }
+    }
+
     /// Return a new [KellyAllocator] with a constraint for maximum permanent loss of capital.
     /// The contents of the original object are moved into the new one. Panics in case a constraint
     /// is already present. Panics in case we already have a constraint of this type.
@@ -58,7 +101,10 @@ impl<'a> KellyAllocator<'a> {
[…]

### 2023-07-13 src/constraints.rs
Change name of constraint from no-short to long-only

@@ -1,3 +1,3 @@
 pub mod capital_loss_constraint;
 pub mod constraint;
-pub mod no_shorting_constraint;
+pub mod long_only_constraint;

### 2023-07-13 src/constraints/long_only_constraint.rs

@@ -2,17 +2,17 @@ use crate::constraints::constraint::{Constraint, InequalityConstraint};
 use crate::model::portfolio::Portfolio;
 use nalgebra::DVector;
 
-/// [NoShortingConstraint] puts a constraint that disallows shorting. It essentially limits the
+/// [LongOnlyConstraint] puts a constraint that essentially disallows shorting. It limits the
 /// fractions such that they are always non-negative.
 #[derive(Debug)]
-pub struct NoShortingConstraint {
-    /// Index representing the fraction it constraints
+pub struct LongOnlyConstraint {
+    /// Index representing the fraction it constraints.
     pub fraction_index: usize,
 }
 
-impl InequalityConstraint for NoShortingConstraint {}
+impl InequalityConstraint for LongOnlyConstraint {}
 
-impl Constraint for NoShortingConstraint {
+impl Constraint for LongOnlyConstraint {
     fn d_constraint_d_fractions(&self, portfolio: &Portfolio) -> DVector<f64> {
         let mut derivative: DVector<f64> = DVector::zeros(portfolio.companies.len());
         derivative[self.fraction_index] = -1.0;

### 2023-07-13 src/kelly_allocation.rs

@@ -12,7 +12,7 @@ use slog::{info, Logger};
 use crate::analysis::{all_outcomes, worst_case_outcome, Outcome};
 use crate::constraints::capital_loss_constraint::CapitalLossConstraint;
 use crate::constraints::constraint::InequalityConstraint;
-use crate::constraints::no_shorting_constraint::NoShortingConstraint;
+use crate::constraints::long_only_constraint::LongOnlyConstraint;
 use crate::model::capital_loss::CapitalLoss;
 use crate::model::company::{Company, TOLERANCE};
 use crate::model::errors::Error;
@@ -48,40 +48,39 @@ impl<'a> KellyAllocator<'a> {
         }
     }
 
-    /// Return a new [KellyAllocator] with a constraint for no shorting, for all company candidates.
-    /// The contents of the original object are moved into the new one. Panics in case a single
-    /// constraint of this type is already present.
-    pub fn with_no_shorting_constraint(self, n_candidates: usize) -> KellyAllocator<'a> {
+    /// Return a new [KellyAllocator] with a long-only constraint (no shorting), for all company
+    /// candidates. The contents of the original object are moved into the new one. Panics in case
+    /// a single constraint of this type is already present.
+    pub fn with_long_only_constraints(self, n_candidates: usize) -> KellyAllocator<'a> {
         info!(
             self.logger,
-            "Setting no shorting constraints for all of the {n_candidates}."
+            "Setting long only constraint for all of the {n_candidates}."
         );
 
         if n_candidates < 1 {
-            panic!("Can't make sense of no shorting constraints because there are {n_candidates} candidates.")
+            panic!("Got {n_candidates} candidates. Can't add long-only constraint.")
         }
 
         // Fractions are always the first set of unknowns in the system.
-        let no_shorting_constraints: Vec<Box<dyn InequalityConstraint>> = (0..n_candidates)
+        let long_only_constraints: Vec<Box<dyn InequalityConstraint>> = (0..n_candidates)
             .map(|i| {
-                Box::new(NoShortingConstraint { fraction_index: i })
-                    as Box<dyn InequalityConstraint>
+                Box::new(LongOnlyConstraint { fraction_index: i }) as Box<dyn InequalityConstraint>
             })
             .collect();
 
         if self
             .inequality_constraints
             .iter()
-            .any(|c| c.type_id() == no_shorting_constraints[0].type_id())
+            .any(|c| c.type_id() == long_only_constraints[0].type_id())
         {
             panic!(
[…]

### 2023-08-18 tests/integration.rs
Application with the CLI

@@ -1,4 +1,4 @@
-use charlie::env::create_test_logger;
+use charlie::env::{create_logger, create_test_logger, get_project_dir};
 use charlie::kelly_allocation::{KellyAllocator, MAX_ITER, SOLVER_TOLERANCE};
 use charlie::model::errors::Error;
 use charlie::model::portfolio::AllocationInput;
@@ -7,119 +7,21 @@ use charlie::utils::assert_close;
 use charlie::validation::result::{Problem, Severity, ValidationResult};
 use charlie::{allocate, analyze, validate};
 use slog::info;
+use slog::Level::Info;
 
 /// Make assertion tolerance the same as the fraction tolerance (no point in more accuracy).
 const ASSERTION_TOLERANCE: f64 = SOLVER_TOLERANCE;
 
-const TEST_YAML: &str = "
-          candidates:
-          - name: A
-            ticker: A
-            description: Business A
-            market_cap: 238.0e9
-            scenarios:
-              - thesis: Unexpected stuff happens
-                intrinsic_value: 0.0
-                probability: 0.05
-              - thesis: Core business keeps losing earnings power
-                intrinsic_value: 170.0e9
-                probability: 0.3
-              - thesis: Business doesn't grow, earnings kept flat
-                intrinsic_value: 270.0e9
-                probability: 0.5
-              - thesis: Earnings grow slightly
-                intrinsic_value: 360.0e9
-                probability: 0.15
-
-          - name: B
-            ticker: B
-            description: Business B
-            market_cap: 363.0e6
-            scenarios:
-              - thesis: Unexpected stuff happens
-                intrinsic_value: 0.0
-                probability: 0.05
-              - thesis: Liquidation value
-                intrinsic_value: 350.0e6
-                probability: 0.5
-              - thesis: Cycle moves upward and the market values the business correctly
-                intrinsic_value: 900.0e6
-                probability: 0.45
-
-          - name: C
-            ticker: C
-            description: Business C
-            market_cap: 35.3e6
-            scenarios:
-              - thesis: >
-                    They don't manage to liquidate and it turns out that they're incompetent as
-                    they were in the past
-                intrinsic_value: 0.0
-                probability: 0.1
-              - thesis: They manage to liquidate at 25% of similar realized prices in the past
-                intrinsic_value: 33.5e6
-                probability: 0.5
[…]

### 2023-08-20 tests/integration.rs
Reorganize integration tests such that they read from file

@@ -1,4 +1,4 @@
-use charlie::env::{create_logger, create_test_logger, get_project_dir};
+use charlie::env::{create_test_logger, get_project_dir};
 use charlie::kelly_allocation::{KellyAllocator, MAX_ITER, SOLVER_TOLERANCE};
 use charlie::model::errors::Error;
 use charlie::model::portfolio::AllocationInput;
@@ -7,21 +7,21 @@ use charlie::utils::assert_close;
 use charlie::validation::result::{Problem, Severity, ValidationResult};
 use charlie::{allocate, analyze, validate};
 use slog::info;
-use slog::Level::Info;
 
 /// Make assertion tolerance the same as the fraction tolerance (no point in more accuracy).
 const ASSERTION_TOLERANCE: f64 = SOLVER_TOLERANCE;
 
 /// Helper function to load the test YAML file as a string.
-fn load_test_input_file() -> String {
-    let test_file_path = get_project_dir().join("tests").join("test_data.yaml");
+fn load_test_file_content(file_name: &str) -> String {
+    let test_file_path = get_project_dir().join("tests").join(file_name);
     std::fs::read_to_string(&test_file_path)
         .expect("Did not manage to read test file in PROJECT_DIR/tests/test_data.yaml.")
 }
 
 #[test]
 fn test_create_candidates_and_validate() {
-    let input: AllocationInput = serde_yaml::from_str(&load_test_input_file()).unwrap();
+    let input: AllocationInput =
+        serde_yaml::from_str(&load_test_file_content("test_data_no_constraints.yaml")).unwrap();
 
     assert_eq!(input.candidates.len(), 6);
 
@@ -100,7 +100,8 @@ fn test_create_candidates_and_validate() {
 #[test]
 fn test_allocate_with_validation_problems() {
     // Change the probability of the first scenario from 0.05 to 0.03
-    let mut input: AllocationInput = serde_yaml::from_str(&load_test_input_file()).unwrap();
+    let mut input: AllocationInput =
+        serde_yaml::from_str(&load_test_file_content("test_data_no_constraints.yaml")).unwrap();
     input.candidates[0].scenarios[0].probability = 0.03;
 
     let logger = create_test_logger();
@@ -124,7 +125,8 @@ fn test_allocate_with_validation_problems() {
 fn test_allocate_with_no_candidates_after_filtering() {
     // Keep only two candidates and change the numbers such that one has negative expected return
     // and the other has no downside scenario
[…]

### 2023-08-21 doc/paper.tex
Add Robi as author
Update the paper with constraints

@@
 \author[$\dagger$]{Vuko Vuk\v{c}evi\'{c}}
+\author[$\dagger$]{Robert Keser}
 
-\affil[$\dagger$]{In silico Ltd., Zagreb, Croatia, vuko.vukcevic@insilico.hr}
+\affil[$\dagger$]{In silico Ltd., Zagreb, Croatia,\linebreak vuko.vukcevic@insilico.hr, robert.keser@insilico.hr}
 %\affil[$\star$]{SimScale GmbH, Munich, Germany, vvukcevic@simscale.com}
@@
 
+The paper attempts to provide a mathematical model and a tool for the focused
+investing strategy as advocated by Buffett~\cite{berkshireLetters},
+Munger~\cite{almanack}, and others from this investment community.
+The approach presented here assumes that the investor's role is to think about
+probabilities of different outcomes for a set of businesses. Based on these
+assumptions, the tool calculates the optimal allocation of capital for
+each of the investment candidates. The model has the option to provide
+constraints that ensure: no shorting, no use of leverage, providing a maximum
+limit to the risk of permanent loss of capital, and maximum individual
+allocation. The software is made available for public use.\\
+
 This paper consist of two parts, one without the mathematics and one with the
@@
 still read the preceding sections since they provide context within which the
-mathematics is built on.
+mathematics is built on. Moreover, for mathematically skilled reader, the
+problem formulated and solved here represents a non--linear constrained
+optimization problem.
 
@@
 cash comes sporadically, and In silico has been a net buyer over the last four
-years: A trend that will likely continue. When the excess cash comes in, the
-following question arises: How much money to put in which stocks? That question
-had been answered by looking at our fundamental analysis for each company, and
-performing some vague hand calculations. That of course works well because of
-all the other uncertainties in the investment process, but we were still
-motivated to:
+years. This is a trend that will likely continue. When the excess cash comes in,
+the following question arises: How much money to put in which stocks? That
+question had been answered by looking at our fundamental analysis for each
+company, and performing some vague hand calculations. That of course works well
+because of all the other uncertainties in the investment process, but we were
+still motivated to:
 \begin{enumerate}
@@
     \item An arbitrary number of scenarios for each company,
+    \item Multiple constraints for modelling our preference towards no shorting,
+    no use of leverage, and providing a maximum value for the risk of permanent
+    loss of capital,
     \item A fundamentals--based analysis (after all, a stock is an ownership
@@
 of competence. And according to Charlie Munger, Warren Buffett, Mohnish Pabrai
-and the like-minded others from whom the author got the inspiration for this
+and the like-minded others from whom the authors got the inspiration for this
 work, one should focus precisely on getting such great businesses for a fair
@@
 form~\footnote{It is on the author's TODO list to try and prove this.}. Here's a
-soft, non-mathematical reasoning on why the author thinks this is fine: The
+soft, non-mathematical reasoning on why the authors think this is fine: The
 assumption is made in order to write the equations in terms of probabilities
@@
     \item No shorting allowed. From a purely mathematical point of view,
-    shorting would be allowed. Without detailed math, it's easy to see how a
+    shorting would be allowed. Without detailed math, it is easy to see how a
     company with a negative expected return would result in a short position.
     However, due to the asymmetry of the potential losses compared to gains,
-    coupled with the usual time--frame limit that comes with shorting, the
-    author decided to avoid it.
+    coupled with the usual time--frame limit that comes with shorting, we
+    provided a constraint for being long--only.
     \item No use of leverage allowed. Again, from a purely mathematical point of
@@
     right. However, if one is absolutely (100\%) sure that a company does not
-    have a downside, then the solution is to lever up as much as possible, and
-    put all the money in that one company. It is our feeling that the best way
-    to handle such cases is outside of this framework. Alternatively, one can
-    always model an unknown downside scenario with a small probability (say
-    5--20\%) and intrinsic value of zero.
+    have a downside, the solution is to put all the money in that one company.
+    It is our feeling that the best way to handle such cases is outside of this
+    framework. Alternatively, one can always model an unknown downside scenario
+    with a small probability (say 5--20\%) and intrinsic value of zero.
 \end{itemize}
 
-These assumptions and margins of safety are embedded into the framework in order
-to try and tie the rational mathematics with common--sense from other
-disciplines.
+These assumptions and margins of safety are embedded into the framework as
+constraints in order to try and tie the rational mathematics with the real
+world.
 
@@
 $N_a$\footnote{The assumption regarding the infinite number of allocations is
-something that the author is slightly uncomfortable with, but \textit{feels} it
-is fine because of the margins of safety embedded into the thinking that goes
+something that the authors are slightly uncomfortable with, but \textit{feels}
+it is fine because of the margins of safety embedded into the thinking that goes
 into assessing each investment opportunity.}, the following relation holds:
@@
 
+\subsection{Constraints}
+\label{sec:constraints}
+
+\noindent Equation~\eqref{eq:8} has no constraints, meaning that after solving
+the system of equations, the resulting fractions $f_j$ might be negative and
+greater than one. This would imply short positions and use of leverage,
+respectively. A general inequality constraint may be written in the following
+form:
+\begin{equation}
+\label{eq:9}
+    I(f_j) \le 0
+\end{equation}
+
+\noindent where, for example, $I(f_j) \equiv -f \le 0$ models a long-only
+constraint that would make sure that the fractions are positive. In order to
+transform the inequality constraint into an equality constraint, we introduce a
+slack variable $s$ that must be positive:
+\begin{equation}
+\label{eq:10}
+    I(f_j) + s = 0 \equiv \mathcal{C}(f_j, s) = 0
+\end{equation}
+
+\noindent where the second part of the equation introduces a useful substitution
+for deriving the constrained system later on.\\
+
+\indent In this work, we define four constraints that serve as additional
+margins of safety in a focused investment approach:
+\begin{enumerate}
+    \item Long--only constraint ensures that a fraction cannot be negative and
+    is added to all candidate companies. Adding this constraint means: "I do not
+    allow short positions.":
+    \begin{equation}
+        \label{eq:11}
+            f \ge 0 \rightarrow -f + s = 0
+        \end{equation}
+
+    \item Maximum leverage constraint ensures that the leverage is limited up to
+    $L$. Note that $L = 0$ implies no leverage. Adding this constraint means: "I
+    want to limit the use of leverage (to a minimal amount).":
+        \begin{equation}
+        \label{eq:12}
+            \sum_j^{N_c} f_j \le 1 + L \rightarrow \sum_j^{N_c} f_j - 1 - L + s = 0
+        \end{equation}
+
+    \item Maximum individual allocation constraint ensures that a fraction does
+    not exceed the specified amount. Adding this constraint means: "I feel
+    uncomfortable putting more than $M$ of my capital into a single company.":
+        \begin{equation}
+        \label{eq:13}
+            f \le M \rightarrow f - M + s = 0
+        \end{equation}
+
+    \item Maximum allowable permanent capital loss constraint ensures that the
+    worst--case outcome does not exceed losing a specified amount of capital
+    with a specified probability. Adding this constraint means: "Under a
+    worst--case scenario, I am comfortable permanently losing $K$ (e.g. 25\%) of
+    my capital with probability $P$ (e.g. 0.1\%).":
+        \begin{equation}
+        \label{eq:14}
+            \sum_j^{N_c} f_j \min(p_{ij} k_{ij}) \ge P \cdot K \rightarrow
+            -\sum_j^{N_c} \left(f_j \min(p_{ij} k_{ij}) \right) + P \cdot K + s = 0
+        \end{equation}
+    \noindent where $\min(p_{ij} k_{ij})$ is the worst--case outcome across all
+    scenarios for the $j$-th candidate company. Here, the minimum returns
+    $k_{ij}$ and the maximum worst--case return $K$ are both negative by
+    convention, indicating a loss, hence the $\ge$ sign.
+
+\indent Note that the maximum allowable permanent capital loss constraint
+given by equation~\eqref{eq:14} only works with the long--only constraint
+because the fraction $f_j$ is assumed positive. It is also important to note
+that here we do not talk about {\it temporary} loss of capital due to short-term
+stock market fluctuations, but rather {\it permanent} loss of capital due to the
+fundamental business environment of candidate companies. In addition, it is of
+course possible for one to lose more than the specified maximum allowable amount
+of capital because of the assumptions made with respect to the inputs: After
+all, coming up with intrinsic values and probabilities for each scenario is more
+of an art than a science.
+
+\end{enumerate}
+
+\subsection{Constrained System: Putting It All Together}
+\label{sec:constrainedSystem}
+
+\noindent The growth function~\eqref{eq:4} can be constrained with an arbitrary
+amount of constraints~\eqref{eq:10} by introducing a Lagrangian:
+\begin{equation}
+\label{eq:15}
+    \mathcal{L}(f_j, \lambda_l)
+  =
+    \mathcal{G}(f_j) - \sum_l^{N_l} \lambda_l \mathcal{C}_l(f_j, s_l)
+\end{equation}
+
+\noindent where $N_l$ is the number of constraints and $l$ denotes the $l$-th
+constraint defined either by $l$-th Lagrange multiplier $\lambda_l$ or the
+$l$-th slack variable $s_l$. There are two necessary conditions for finding a
+constrained maximum of the growth function:
+\begin{equation}
+\label{eq:16}
+    \alpha_i(f_j, \lambda_l, s_l)
+  =
+    \frac{\partial \mathcal{L}(f_j, \lambda_l)}{\partial f_j}
+  =
+    \frac{\partial \mathcal{G}(f_j)}{\partial f_j}
+  - \sum_l^{N_l} \lambda_l \frac{\partial{\mathcal{C}_l(f_j, s_l)}}{\partial f_j}
+  =
+    0
+\end{equation}
+
+\begin{equation}
+\label{eq:17}
+    \beta_i(f_j, s_l)
+  =
+    \frac{\partial \mathcal{L}(f_j, \lambda_l)}{\partial \lambda_l}
+  =
+    - \mathcal{C}_l(f_j, s_l)
+  =
+    0
+\end{equation}
+
+\noindent where $\alpha_i$ and $\beta_i$ have been introduced as a shorthand
+notation that distinguishes between two vector equations: There are $N_c$
+of $\alpha$ equations and $N_l$ of $\beta$ equations. Therefore, there are
+$N_c + N_l$ equations, but $N_c + 2N_l$ unknowns, because of $N_c$ unknown
+fractions, $N_l$ unknown Lagrange multipliers $\lambda_l$ and $N_l$ unknown
+slack variables $s_l$. However, an inequality constraint cannot be active and
+inactive at the same time. An active constraint is characterized by $\lambda_l
+\neq 0$ and $s_l = 0$, whereas an inactive constraint is characterized by
+$\lambda_l = 0$ and $s_l > 0 \ne 0$. This means that we have to solve $2^{N_l}$
+nonlinear systems to cover all combinations of constraints and pick the best
+solution. As an example, adding all constraints mentioned
+in~\autoref{sec:constraints} for a portfolio with five candidate companies would
+result in having to solve $2^{12} = 4 096$ systems, while having ten candidate
+companies would imply solving $2^{22} = 4 194 304$ systems, demonstrating the
+exponential complexity of the problem.
+
 \section{Numerics}
@@
 
-\noindent The equation~\eqref{eq:8} can be written succinctly as:
+\noindent The equations~\eqref{eq:16} and~\eqref{eq:17} can be written
+succinctly as:
 \begin{equation}
-\label{eq:9}
-    \mathcal{F}_i(f_j) = 0
+\label{eq:18}
+    \mathcal{F}_i(x_i) = 0
 \end{equation}
 
-\noindent Because $\mathcal{F}_i$ is a non--linear equation in $f_j$,
+\noindent where $x_i$ is a vector of unknown fractions and unknown Lagrange
+multipliers or slack variables:
+\begin{equation}
+\label{eq:19}
+    x_i
+  =
+    \{ f_1, f_2, \ldots, f_{N_c}, \lambda_1 | s_1, \lambda_2 | s_2,
+    \ldots, \lambda_{N_l} | s_{N_l} \}
+\end{equation}
+
+\indent Because $\mathcal{F}_i$ is a non--linear equation in $f_j$,
 the Newton--Raphson method is used to find a numerical solution. The method is
@@
 \begin{equation}
-\label{eq:10}
-    \mathcal{F}_i^o + \sum_{j}^{N_c} \mathcal{J}_{ij}^o (f_j^n - f_j^o) = 0
+\label{eq:20}
+    \mathcal{F}_i^o + \sum_{i}^{N_c + N_l} \mathcal{J}_{ij}^o (x_j^n - x_j^o) = 0
 \end{equation}
 
-\noindent where $\mathcal{J}_{ij}$ is the Jacobian of $\mathcal{F_i}$, and
+\noindent where $\mathcal{J}_{ij}$ is the Jacobian of $\mathcal{F}_i$, and
 superscripts $^n$ and $^o$ denote the new value and the old value, respectively.
-The Jacobian is:
+
+\subsection{Jacobian for an Active Constraint}
+\label{sec:jacobianActive}
+
+\noindent If a constraint is active ($\lambda_l \ne 0, s_l = 0$), the
+corresponding unknown is the Lagrange multiplier and the Jacobian has the
+following form:
 \begin{equation}
-\label{eq:11}
+\label{eq:21}
     \mathcal{J}_{ij}
+  =
+    \begin{pmatrix}
+        \frac{\partial \alpha_i}{\partial f_j}
+      & \frac{\partial \alpha_i}{\partial \lambda_j} \\
+        \frac{\partial \beta_i}{\partial f_j}
+      & \frac{\partial \beta_i}{\partial \lambda_j} \\
+    \end{pmatrix}
+  =
+    \begin{pmatrix}
+        \frac{\partial^2 \mathcal{G}}{\partial f_i \partial f_j}
+      - \sum_j^{N_l} \left( \lambda_j \frac{\partial^2 \mathcal{C}_i}{\partial f_j^2} \right)
+      & -\frac{\partial \mathcal{C}_i}{\partial f_j} \\
+        -\frac{\partial \mathcal{C}_i}{\partial f_j}
+      & 0 \\
+    \end{pmatrix}
+\end{equation}
+
+\subsection{Jacobian for an Inactive Constraint}
+\label{sec:jacobianInactive}
+
+\noindent If a constraint is inactive ($\lambda_l = 0, s_l > 0 \ne 0$), the
+corresponding unknown is the slack variable and the Jacobian has the following
+form:
+\begin{equation}
+\label{eq:22}
+    \mathcal{J}_{ij}
+  =
+    \begin{pmatrix}
+        \frac{\partial \alpha_i}{\partial f_j}
+      & \frac{\partial \alpha_i}{\partial s_j} \\
+        \frac{\partial \beta_i}{\partial f_j}
+      & \frac{\partial \beta_i}{\partial s_j} \\
+    \end{pmatrix}
+  =
+    \begin{pmatrix}
+        \frac{\partial^2 \mathcal{G}}{\partial f_i \partial f_j}
+      & 0 \\
+        -\frac{\partial \mathcal{C}_i}{\partial f_j}
+      & -1 \\
+    \end{pmatrix}
+\end{equation}
+
+\noindent Partial derivatives of the constraint function can be readily obtained
+because all constraints presented in this work are simple analytical functions
+(see~\autoref{sec:constraints}). For completeness, the Hessian of the growth
+function that appears in the upper--right corner is:
+\begin{equation}
+\label{eq:23}
+    \frac{\partial^2 \mathcal{G}}{\partial f_i \partial f_j}
   = 
@@
 \noindent where subscript for outcome $i$ has been changed to $o$ in order to be
-able to write the Jacobian in the standard $ij$ notation.
+able to write the Hessian in the standard $ij$ notation.
 
-The numerical procedure starts by assuming the uniform allocation across
-all companies, i.e. $f_j = f = \frac{1}{N_c}$. Based on the $f_j^o$ in the
-current iteration, the linear system in equation~\eqref{eq:10} is solved to find
-the new solution $f_j^n$. The process is repeated until sufficient level of
-accuracy is reached~\footnote{Usually two significant digits is enough
-considering the uncertainty and the judgment required for the inputs.}.
+\indent Note that the equations~\eqref{eq:21} and~\eqref{eq:22} are presented in
+a way that all constraints are either active or inactive. Since we have to solve
+$2^{N_l}$ systems characterized by active/inactive status of each constraint,
+that simply means that we need to add respective active/inactive contributions
+to the Jacobian $\mathcal{J}_{ij}$ and the right--hand--side $\mathcal{F}_i^o$
+in equation~\eqref{eq:20} for a particular constraint $j$.
 
-A practical approach is used to prevent shorting and use of leverage. Shorting
-would occur if any of the fractions is negative, and is prevented by filtering
-out the candidate companies with negative expected return before solving the
-equations. The use of leverage is avoided by normalizing the fractions after
-obtaining a numerical solution, if the sum of all fractions exceeds 1. The
-author is well--aware that this is somewhat a brute--force approach, and it will
-be a topic of future work to avoid shorting and use of leverage by using
-inequality constraints.
+\subsection{Notes on the Numerical Procedure}
+\label{sec:numericalNotes}
+
+\noindent The numerical procedure starts by assuming the uniform allocation
+across all companies, i.e. $f_j = f = \frac{1}{N_c}$. Based on the $f_j^o$ in
+the current iteration, the linear system in equation~\eqref{eq:10} is solved to
+find the new solution $f_j^n$. The process is repeated until sufficient level of
+accuracy is reached.\\
+
+\indent After solving all $2^N_l$ solutions, we end up with less than $2^N_l$
+viable solutions. A solution is considered viable if the Newton--Raphson
+procedure managed to find a numerical solution and if all slack variables
+for all inactive constraints are positive. Finding the best solution out of all
+viable solutions would require evaluating the growth function for all
+solutions, which is particularly challenging due to the product series
+in equation~\eqref{eq:3}. In this work, we take a pragmatic approach and
+simply choose the solution that maximizes the expected value of the portfolio
+out of a set of most diversified solutions, i.e. the solutions with the highest
+number of non--zero allocations.
+
+\section{Basic Validation Tests}
+\label{sec:validation}
+
+\noindent In order to validate the numerical model, a basic example of five
+candidate companies is considered, where each candidate has the same set of
+scenarios (probabilities and intrinsic values) and the same market cap.
+The inputs that present a 50\% loss and 100\% gain with 50--50 probabilities are
+presented in~\autoref{tab:validationCompany}.\\
+
+\begin{table}[!ht]
+\caption{Company for a validation test with a market cap of 1.}
+\vspace{0.25cm}
+\centering
+\begin{tabular}{l|c|c}
+Scenario & Intrinsic value & Probability \\
+\hline
+50\% down with 50\% probability & 0.5 & 50\% \\
+100\% up with 50\% probability & 2 & 50\% \\
+\end{tabular}%
+\label{tab:validationCompany}%
+\end{table}%
+
+\indent Solving the system without any constraints yields a uniform allocation
+of 35\% of capital in each company. Note that because we considered 5 companies,
+that implies 75\% leverage ($5 \cdot 35\% = 175\%$). Even without the
+constraint for maximum allocation of capital, just maximizing the long--term
+growth--rate of assets prefers a diversified solution, which is expected.\\
+
+\indent Adding a constraint for no leverage (given by equation~\eqref{eq:12} and
+setting $L = 0$), results in a uniform allocation of 20\%, as expected. It is
+straightforward to show that the worst--case outcome in such a portfolio implies
+permanently losing 50\% of the capital with probability of 3.125\%.\\
+
+\indent The final test is done by setting the maximum allowable permanent
+capital loss constraint as given by equation~\eqref{eq:14}. Setting $P = 5\%$
+and $K = 50\%$, indicating that we are comfortable risking to lose 50\% of the
+capital with 5\% probability, results in a uniform allocation of 2\%. Because
+only 10\% of capital is invested in that case, there is a possibility of losing
+5\% of capital with probability of 3.125\%. In the worst--case scenario, the
+probability--weighted return is $-0.5 \cdot 0.02 \cdot 0.5 \cdot 5 = -2.5\%$,
+which is equal to $P \cdot K$.
 
@@
 
-\noindent As an example, consider seven candidate companies, each with two to
-four scenarios. Each scenario is represented by an intrinsic value and the
-probability of the scenario happening (or intrinsic value being reached at some
-point in the future). Note that how these numbers are obtained is outside of the
-scope of this work, although it is important to stress that the validity and
-conservative assumptions behind these numbers are probably the most important
-part of an investor's job. The example inputs are presented
-in~\autoref{tab:companyA} to~\autoref{tab:companyF}.
+\noindent With the basic validation of the numerical model performed, the
+attention is moved to a realistic example. Consider five candidate companies,
+each with up to three scenarios. Each scenario is represented by an intrinsic
+value and the probability of the scenario happening (or intrinsic value being
+reached at some point in the future). Note that how these numbers are obtained
+is outside of the scope of this work, although it is important to stress that
+the validity and conservative assumptions behind these numbers are probably the
+most important part of an investor's job. The example inputs are presented
+in~\autoref{tab:companyA} to~\autoref{tab:companyE}.
 
 \begin{table}[!ht]
-\caption{Company A with current market cap of 214B USD.}
+\caption{Company A with current market cap of 225B USD.}
 \vspace{0.25cm}
@@
 Total loss & 0 USD & 5\% \\
-Bear thesis & 170B USD & 30\% \\
-Base thesis & 270B USD & 50\% \\
-Bull thesis & 360B USD & 15\% \\
+Base thesis & 270B USD & 60\% \\
+Bull thesis & 420B USD & 35\% \\
 \end{tabular}%
@@
 \begin{table}[!ht]
-\caption{Company B with current market cap of 306M USD.}
+\caption{Company B with current market cap of 450M USD.}
 \vspace{0.25cm}
@@
 Total loss& 0 USD & 5\% \\
-Bear thesis & 300M USD & 50\% \\
+Bear thesis & 350M USD & 50\% \\
 Base thesis & 900M USD & 45\% \\
@@
 \begin{table}[!ht]
-\caption{Company C with current market cap of 34M GBP.}
+\caption{Company C with current market cap of 39M GBP.}
 \vspace{0.25cm}
@@
 Total loss & 0 GBP & 10\% \\
-Bear thesis & 33.5M GBP & 50\% \\
-Base thesis & 135M GBP & 40\% \\
+Bear thesis & 34M GBP & 40\% \\
+Base thesis & 135M GBP & 50\% \\
 \end{tabular}%
@@
 \begin{table}[!ht]
-\caption{Company D with current market cap of 806M SGD.}
+\caption{Company D with current market cap of 751M SGD.}
 \vspace{0.25cm}
@@
 \hline
-Bear thesis & 330M SGD & 40\% \\
-Base thesis & 1B SGD & 50\% \\
-Bull thesis & 4B SGD & 10\% \\
+Bear thesis & 330M SGD & 30\% \\
+Base thesis & 1B SGD & 70\% \\
 \end{tabular}%
@@
 \begin{table}[!ht]
-\caption{Company E with current market cap of 17.6M CAD.}
+\caption{Company E with current market cap of 126B HKD.}
 \vspace{0.25cm}
@@
 \hline
-Total loss & 0 CAD & 5\% \\
-Bear thesis & 10M CAD & 25\% \\
-Base thesis & 25M CAD & 70\% \\
+Total loss & 0 HKD & 5\% \\
+Bear thesis & 50B HKD & 10\% \\
+Base thesis & 300B HKD & 85\% \\
 \end{tabular}%
@@
 
-\begin{table}[!ht]
-\caption{Company F with current market cap of 581M USD.}
-\vspace{0.25cm}
-\centering
-\begin{tabular}{l|c|c}
-Scenario & Intrinsic value & Probability \\
-\hline
-Total loss & 0 USD & 20\% \\
-Base thesis & 1B USD & 80\% \\
-\end{tabular}%
-\label{tab:companyF}%
-\end{table}%
-
-\noindent Based on the inputs, the portfolio allocation that maximizes the
-long--term growth--rate of capital is presented in~\autoref{tab:results}.
+\noindent Based on these inputs, with the long--only strategy, without leverage,
+and with maximum individual allocation of 30\%, the portfolio allocation that
+maximizes the long--term growth--rate of capital is presented
+in~\autoref{tab:results}.
 
@@
 \begin{tabular}{l|c|c|c|c|c|c}
-Company & A & B & C & D & E & F \\
+Company & A & B & C & D & E \\
 \hline
-Allocation fractions & 13\% & 27\% & 26\% & 12\% & 12\% & 10\% \\
+Allocation fractions & 30\% & 8\% & 30\% & 2\% & 30\% \\
 \end{tabular}%
@@
 \begin{itemize}
-    \item Expected gain of 61 cents for every dollar invested,
-    \item Cumulative probability of loss of capital of 13\%,
-    \item Permanent loss of 95\% of capital with probability of 0.001\%.
+    \item Expected gain of 32 cents for every dollar invested,
+    \item Cumulative probability of loss of capital of 16\%,
+    \item Permanent loss of 60\% of capital with probability of 0.008\%.
 \end{itemize}
 
-\noindent The last item is particularly interesting to the author. According to
-Actuarial Life Tables in~\cite{lifeTables}, the probability of the (currently 33
-year--old) author dying within the next year is approximately 0.25\%. That is
-two orders of magnitude higher than the probability of the permanent loss of
-capital for this portfolio. Considering that the portfolio has six stocks, that
-is a very strong argument against excessive diversification, especially if:
+\noindent The last item is particularly interesting to the authors. According to
+Actuarial Life Tables in~\cite{lifeTables}, the probability of the (currently 34
+year--old, the oldest) author dying within the next year is approximately
+0.26\%\footnote{This is the only thing that's probably worth measuring in the
+basis points.}. That is two orders of magnitude higher than the probability of
+the permanent loss of capital for this portfolio. Considering that the portfolio
+has five stocks, that is a very strong argument against excessive
+diversification, especially if:
 \begin{itemize}
@@
 \noindent The observation about excessive diversification is inline with the
-thoughts from the Poor Charlie's Almanack~\cite{almanack}.
+thoughts from the Poor Charlie's Almanack~\cite{almanack} and one of the
+lectures from Li Lu that the authors frequently watch~\cite{liLuLecture}.
 
-\section{Future Work}
+\section{Problems, Discussion, and Future Work}
 \label{sec:futureWork}
 
-\noindent The future work will focus on adding a constraint for the maximum
-allowable risk of the permanent loss of capital. The idea is that an investor
-could input the probability of the permanent loss of capital and the fraction of
-capital one is willing to risk, and the allocation will change accordingly. For
-example, an investor might be fine with losing 50\% of capital with the
-probability of 0.1\% as a worst--case outcome, which should be an additional
-input to the system.
+\noindent There are a couple of technical problems that the authors observed:
+\begin{enumerate}
+    \item It is possible that a given non--linear system for a particular
+    combination of constraint statuses (active/inactive) does not converge. This
+    may happen if the resulting matrix is singular, or if the Newton--Raphson
+    algorithm does not find the solution within a prescribed number of
+    iterations. These errors are ignored, which means that it is possible that
+    the real best solution is not found due to numerical issues.
+    \item The exponential complexity of the model makes it challenging to
+    use for more than several candidate companies without using significant
+    compute resources. For example, having a 20 candidate companies with all
+    constraints would result in around 4 trillion non--linear systems to solve.
+    Therefore, the model is not suitable for cases with excessive
+    diversification, although there is a possibility to filter some of these
+    upfront without attempting to solve them, which may be one of the topics for
+    future work.
+\end{enumerate}
+
+\indent To conclude, the authors believe that the most challenging aspect of an
+investor's work that might use this software is to think hard about the inputs.
+The authors see the usefulness of this software mainly in:
+\begin{itemize}
+    \item Forcing the investors to think consistently in terms of probabilities
+    and long--term business outcomes across the range of candidate companies,
+    \item Calculating the optimal allocation in a short amount of time, based on
+    the investor's inputs.
+\end{itemize}
+
+\section{Access to the Software}
 

### 2023-09-30 src/kelly_allocation.rs
Prefer the most diversified solutions when picking the winner

@@ -1,6 +1,7 @@
 use bitvec::order::Lsb0;
 use bitvec::slice::BitSlice;
 use bitvec::view::BitView;
+use itertools::Itertools;
 use nalgebra::{DMatrix, DVector};
 use num_traits::pow;
 use num_traits::pow::Pow;
@@ -349,7 +350,7 @@ impl<'a> KellyAllocator<'a> {
                             "Solution is not viable, skipping it. Solution vector: {x}."
                         );
                     } else {
-                        info!(
+                        debug!(
                             self.logger,
                             "This is a viable solution. Adding it to the list of all solutions. \
                             Solution vector: {x}."
@@ -367,15 +368,54 @@ impl<'a> KellyAllocator<'a> {
             }
         });
 
-        // Assume that the best solution is the one with the highest expected value. This is a poor
-        // man's proxy for choosing the best solution. TODO. Improve
+        // Fail if there are no solutions
+        info!(self.logger, "Found {} viable solutions.", solutions.len());
+        if solutions.is_empty() {
+            return Err(Error {
+                code: "did-not-find-a-single-viable-solution".to_string(),
+                message: format!(
+                    "Did not manage to find a single viable numerical solution. \
+                         This may happen for multiple reasons. Check whether the input data would \
+                         suggest a very strong bias towards a single/few investments. Check whether \
+                         the constraints are too strict.\n\
+                         Errors in individual solutions are {}:", all_error_strings
+                ),
+            });
+        }
+
+        // The best solution is considered to be the one that has a highest expected return among
+        // a set of the most diversified solutions. Most diversified solutions are the ones who have
+        // the highest number of non-zero allocations. This is a poor man's proxy of finding the
+        // best solution, which should be done by evaluating the growth function for each solution.
+        // That would be both tricky and fairly expensive. TODO: Improve.
+        info!(self.logger, "Finding a set of most diversified solutions.");
+        let n_non_zero_allocations_for_solution = solutions
+            .iter()
+            .map(|x| {
+                x.iter()
+                    .enumerate()
+                    .filter(|(i, f)| i < &n_companies && f.abs() > SOLVER_TOLERANCE)
+                    .collect_vec()
+                    .len()
+            })
+            .collect_vec();
+
+        let max_n_non_zero_allocations = n_non_zero_allocations_for_solution.iter().max().unwrap();
+        let most_diversified_solutions = solutions
+            .iter()
+            .zip(n_non_zero_allocations_for_solution.iter())
[…]

### 2023-09-30 tests/integration.rs

@@ -1,4 +1,4 @@
-use charlie::env::{create_logger, create_test_logger, get_project_dir};
+use charlie::env::{create_test_logger, get_project_dir};
 use charlie::kelly_allocation::{KellyAllocator, MAX_ITER, SOLVER_TOLERANCE};
 use charlie::model::capital_loss::CapitalLoss;
 use charlie::model::company::Company;
@@ -11,7 +11,6 @@ use charlie::validation::result::{Problem, Severity, ValidationResult};
 use charlie::{allocate, analyze, validate};
 use itertools::Itertools;
 use slog::info;
-use slog::Level::Info;
 
 /// Make assertion tolerance the same as the fraction tolerance (no point in more accuracy).
 const ASSERTION_TOLERANCE: f64 = SOLVER_TOLERANCE;
@@ -302,7 +301,7 @@ fn test_allocate_all_same_with_no_leverage() {
 /// constraint produces a uniform allocation where not all assets are invested.
 #[test]
 fn test_allocate_all_same_with_no_leverage_and_maximum_capital_loss_constraint() {
-    let logger = create_logger(Info);
+    let logger = create_test_logger();
     let input: AllocationInput = create_five_same_candidates(
         Some(true),
         Some(CapitalLoss {

### 2023-10-03 doc/paper.tex
Minor updates to the paper

@@
     loss of capital,
-    \item A fundamentals--based analysis (after all, a stock is an ownership
-    share of business~\cite{intelligentInvestor}).
+    \item A fundamentals--based analysis with a very long time horizon. After
+    all, a stock is an ownership share of business~\cite{intelligentInvestor},
+    in which time--scales are significantly larger than the time--scales of
+    day--to--day stock price fluctuations.
 \end{enumerate}
@@
 
-The underlying philosophy is that one should spend the majority of his time
+The underlying philosophy is that one should spend the majority of their time
 analyzing investments and thinking about intrinsic values under different
@@
     think about what can go wrong, as opposed to dreaming about what can go
-    right. However, if one is absolutely (100\%) sure that a company does not
-    have a downside, the solution is to put all the money in that one company.
-    It is our feeling that the best way to handle such cases is outside of this
+    right. This is in--line with thoughts by Prasad~\cite{darwinInvesting} about
+    the importance of avoiding the errors of investing in a bad company.
+    However, if one is absolutely (100\%) sure that a company does not have a
+    downside, the solution is to put all the money in that one company.  It is
+    our feeling that the best way to handle such cases is outside of this
     framework. Alternatively, one can always model an unknown downside scenario
@@
 $N_a$\footnote{The assumption regarding the infinite number of allocations is
-something that the authors are slightly uncomfortable with, but \textit{feels}
+something that the authors are slightly uncomfortable with, but \textit{feel}
 it is fine because of the margins of safety embedded into the thinking that goes
@@
 the validity and conservative assumptions behind these numbers are probably the
-most important part of an investor's job. The example inputs are presented
-in~\autoref{tab:companyA} to~\autoref{tab:companyE}.
+most important part of an investor's job who is taking a fundamental analysis
+approach. The example inputs are presented in~\autoref{tab:companyA}
+to~\autoref{tab:companyE}.
 
@@
 0.26\%\footnote{This is the only thing that's probably worth measuring in the
-basis points.}. That is two orders of magnitude higher than the probability of
-the permanent loss of capital for this portfolio. Considering that the portfolio
-has five stocks, that is a very strong argument against excessive
-diversification, especially if:
+basis points in this framework.}. That is two orders of magnitude higher than
+the probability of the permanent loss of capital for this portfolio. Considering
+that the portfolio has five stocks, that is a very strong argument against
+excessive diversification, especially if:
 \begin{itemize}
@@
 
+\noindent The software is freely available at TODO. Reader is referred to its
+documentation for instructions on how to install and use the software.
+
 \clearpage

### 2024-02-11 doc/paper.tex

@@
 
-\noindent The software is freely available at TODO. Reader is referred to its
-documentation for instructions on how to install and use the software.
+\noindent The software is freely available as a Rust crate
+at~\url{https://crates.io/crates/charlie}, and its source code is hosted on
+GitLab at~\url{https://gitlab.com/in-silico-public/charlie}. The reader is
+referred to its documentation for instructions on how to install and use the
+software.
 

### 2024-02-23 doc/formal/paper.tex

--- 2024-02-11 doc/paper.tex
@@
 \title{Sizing the bets in a focused portfolio}
-%\title{Capital allocation for a focused portfolio based on fundamental analysis}
+%\title{Capital Allocation With Kelly Criterion for a Focused Portfolio Based On Fundamental Analysis}
 
@@
 
-The paper attempts to provide a mathematical model and a tool for the focused
+The paper provides a mathematical model and a tool for the focused
 investing strategy as advocated by Buffett~\cite{berkshireLetters},
@@
 assumptions, the tool calculates the optimal allocation of capital for
-each of the investment candidates. The model has the option to provide
-constraints that ensure: no shorting, no use of leverage, providing a maximum
-limit to the risk of permanent loss of capital, and maximum individual
-allocation. The software is made available for public use.\\
-
-This paper consist of two parts, one without the mathematics and one with the
-mathematics. For a non--mathematical reader, it is best to skip the two sections
-related to mathematics and numerics, namely~\autoref{sec:mathematics}
-and~\autoref{sec:numerics}. The reader who is interested in mathematics should
-still read the preceding sections since they provide context within which the
-mathematics is built on. Moreover, for mathematically skilled reader, the
-problem formulated and solved here represents a non--linear constrained
-optimization problem.
-
-\section{Motivation}
-\label{sec:motivation}
-
-Before going on to explain the motivation about this work, a bit of context
-might come useful. In silico is an employee--owned engineering consultancy
-company that invests its excess cash into publicly traded stocks. The excess
-cash comes sporadically, and In silico has been a net buyer over the last four
-years. This is a trend that will likely continue. When the excess cash comes in,
-the following question arises: How much money to put in which stocks? That
-question had been answered by looking at our fundamental analysis for each
-company, and performing some vague hand calculations. That of course works well
-because of all the other uncertainties in the investment process, but we were
-still motivated to:
-\begin{enumerate}
-    \item Improve the odds of us behaving rationally (i.e. minimize
-    psychological misjudgments such as anchoring bias~\cite{kahneman},
-    consistency and commitment tendency~\cite{cialdini}, and
-    others with their combined effects~\cite{almanack}),
-    \item Save time (i.e. avoid doing hand calculations).
-\end{enumerate}
-
-Since mathematics is usually good at keeping people rational, and software is
-great for saving time, a decision has been made to write a software that answers
-the following question: For a set of candidate companies and their current
-market capitalization, each having a set of scenarios defined by a probability
-and intrinsic value estimate, how much of our capital to invest in each?
+each of the investment candidates. The model is based on a generalized Kelly
+Criterion with options to provide constraints that ensure: no shorting, limited
+use of leverage, providing a maximum limit to the risk of permanent loss of
+capital, and maximum individual allocation. The software is applied to an
+example portfolio from which certain observations about excessive
+diversification are obtained. In addition, the software is made available for
+public use.\\
 
@@
 
-The answer to this question has been given by Kelly with his widely known Kelly
-formula (sometimes called Kelly criterion)~\cite{kelly}. The Kelly's approach
-starts by maximizing long--term growth of capital when one is presented with
-an infinite amount of opportunities to bet on. This work extends that idea by
-considering:
+A stock is an ownership share of business~\cite{intelligentInvestor}, in
+which time--scales are significantly greater than the time--scales of
+day--to--day stock price fluctuations. The work therefore assumes a
+fundamentals--based investment analysis with a very long time horizon. There are
+generally two parts of a fundamentals--based investment process:
 \begin{enumerate}
-    \item Multiple companies (stocks) in parallel,
-    \item An arbitrary number of scenarios for each company,
-    \item Multiple constraints for modelling our preference towards no shorting,
-    no use of leverage, and providing a maximum value for the risk of permanent
-    loss of capital,
-    \item A fundamentals--based analysis with a very long time horizon. After
-    all, a stock is an ownership share of business~\cite{intelligentInvestor},
-    in which time--scales are significantly larger than the time--scales of
-    day--to--day stock price fluctuations.
+    \item Finding businesses that are investment candidates,
+    \item Deciding how much capital to allocate in each of them.
+\end{enumerate}
+
+\noindent This work does not deal with the first part related to finding
+relevant investment candidates. It deals only with the second part of the
+process, which can be mathematically well formulated, in an attempt to minimize
+the effects of the psychological misjudgements such as anchoring
+bias~\cite{kahneman}, consistency and commitment tendency~\cite{cialdini}, and
+others with their combined effects~\cite{almanack}). Essentially, the question
+to answer is: For a set of candidate companies and their current market
+capitalization, each having a set of scenarios defined by a probability and
+intrinsic value estimate, how much of capital to invest in each?\\
+
+\indent The answer to this question has been given by Kelly~\cite{kelly} with
+his widely known Kelly Criterion. Despite the fact that the Kelly Criterion has
+been published in 1956 and dealt with information theory and horse racing, its
+use is now widespread in the field of finance, with a plethora of recent
+research. For example, Pinelis and Ruppret~\cite{pinelisRuppert} noted that the
+asset allocation fractions (weights) of the machine--learning powered Maximally
+Predictable Portfolio (MPP) are similar to weighting by the Kelly Criterion.
+Meister~\cite{meister} used the Kelly Criterion to demonstrate the advantage of
+an investor who bets on multiple simultaneous investments compared to an
+investor who bets on a single investment. In the limit of infinite number of
+simultaneous and similar bets, the first investor has all the capital invested
+in a diversified fashion, providing a clear advantage. Byrnes and
+Barnett~\cite{byrnesBarnett} went further to generalize the Kelly Criterion with
+a continuous probability distribution modeling the price movement of the stock,
+while Lototsky and Pollok~\cite{lototskyPollok} rigorously analyzed the
+high--frequency limit of the model, concluding that a high--frequency
+strategy can lead to a more aggressive betting strategy.  Proskurnikov and
+Barmish~\cite{proskurnikovBarmish} investigated the possibility of having a
+non--linear control for the Kelly betting, assuming a bounded set for the
+input probabilities instead of assuming that they are known in advance.  The
+Kelly Criterion was recently even extended to the quantum realm by Meister
+and Price~\cite{meisterPrice}, where in addition to the uncertainty related
+to flipping a coin, there is an uncertainty in measurement. Perhaps the most
+related reference to this work is the paper by Jim\'{e}nez et
+al.~\cite{jimenezEtAl} where they extended the Kelly Criterion to both
+multivariate and simultaneous bets and applied it to sports betting in the
+English Premier League. This work largely follows the approach by
+Jim\'{e}nez et al.~\cite{jimenezEtAl}, but applied to a fundamentals based,
+focused investing with the following characteristics:
+\begin{enumerate}
+    \item Multiple companies (stocks) are considered simultaneously,
+    \item Each with an arbitrary number of scenarios,
+    \item Multiple constraints for modeling authors' preference towards no
+    shorting, limited use of leverage, and providing a maximum value for the
+    risk of permanent loss of capital,
+    \item A fundamentals--based analysis with a very long time horizon, without
+    necessarily knowing when the market price will correspond to the
+    fundamentals.
 \end{enumerate}
@@
 next two subsections (\autoref{sec:assumptions} and \autoref{sec:safetyMargin}),
-a discussion on assumptions in a non--mathematical way, and the margin of
-safety, are presented.
+a discussion on assumptions and the margin of safety, are presented.
 
-\subsection{Disclaimers and Assumptions}
+\subsection{Assumptions}
 \label{sec:assumptions}
 
-The underlying philosophy is that one should spend the majority of their time
-analyzing investments and thinking about intrinsic values under different
-scenarios that might play out, independently of outside thoughts and events.
-However, calculating intrinsic value of a company is more of an art than a
-science, especially for a high-quality, growing businesses within one's circle
-of competence. And according to Charlie Munger, Warren Buffett, Mohnish Pabrai
-and the like-minded others from whom the authors got the inspiration for this
-work, one should focus precisely on getting such great businesses for a fair
-price. That means that one shouldn't take what this approach says at face value,
-and one should probably use its guidance infrequently.
+The underlying philosophy with the fundamentals--based focused investing is that
+one should spend the majority of their time analyzing investments and thinking
+about intrinsic values under different scenarios that might play out. However,
+calculating intrinsic value of a company is a process subject to multiple levels
+of uncertainty. Therefore, the reader should always keep in mind that if the
+inputs to the mathematical model are unrealistic, the results will be
+unrealistic as well.
 
@@
 assumption is made that the number of bets is very high (tends to infinity).
-This work does not try to justify this assumption in a strong mathematical
-form~\footnote{It is on the author's TODO list to try and prove this.}. Here's a
-soft, non-mathematical reasoning on why the authors think this is fine: The
-assumption is made in order to write the equations in terms of probabilities
+The assumption is made in order to write the equations in terms of probabilities
 instead of the number of outcomes divided by the number of bets. Therefore, as
@@
     \item No shorting allowed. From a purely mathematical point of view,
-    shorting would be allowed. Without detailed math, it is easy to see how a
-    company with a negative expected return would result in a short position.
-    However, due to the asymmetry of the potential losses compared to gains,
-    coupled with the usual time--frame limit that comes with shorting, we
+    shorting would be allowed. Without detailed mathematics, it is easy to see
+    how a company with a negative expected return would result in a short
+    position. However, due to the asymmetry of the potential losses compared to
+    gains, coupled with the usual time--frame limit that comes with shorting, we
     provided a constraint for being long--only.
-    \item No use of leverage allowed. Again, from a purely mathematical point of
-    view, use of leverage would sometimes be useful. The thinking in avoiding to
-    use leverage is that no--one should be in a hurry to get rich, and should
-    avoid risking good night's sleep based on short--term market fluctuations,
-    which are fairly hard to predict consistently (unless you work at
-    Renaissance Technologies, in which case you can scratch this).
+    \item Limited use of leverage. Although leverage is especially useful when
+    dealing with high--conviction ideas, this constraint is introduced to model
+    authors' conservative behavior.
     \item No company without at least one downside scenario is allowed. By
-    disallowing inputs without downside, the framework forces you to focus and
-    think about what can go wrong, as opposed to dreaming about what can go
+    disallowing inputs without downside, the framework forces us to focus and
+    think about what can go wrong, as opposed to focusing on what can go
     right. This is in--line with thoughts by Prasad~\cite{darwinInvesting} about
@@
     However, if one is absolutely (100\%) sure that a company does not have a
-    downside, the solution is to put all the money in that one company.  It is
-    our feeling that the best way to handle such cases is outside of this
-    framework. Alternatively, one can always model an unknown downside scenario
-    with a small probability (say 5--20\%) and intrinsic value of zero.
+    downside, the solution is to put a significant amount of leveraged capital
+    in that one company. There are two options to handle such cases: i) Either
+    outside of this framework or ii) Modeling an unknown downside scenario with
+    a small probability and intrinsic value of zero.
 \end{itemize}
@@
 These assumptions and margins of safety are embedded into the framework as
-constraints in order to try and tie the rational mathematics with the real
-world.
+constraints in order to try and tie the mathematical model and the uncertainty
+around inputs.
 
@@
 $N_a$\footnote{The assumption regarding the infinite number of allocations is
-something that the authors are slightly uncomfortable with, but \textit{feel}
-it is fine because of the margins of safety embedded into the thinking that goes
-into assessing each investment opportunity.}, the following relation holds:
+somewhat incompatible with a focused investment strategy where a small number of
+stocks are considered at any point in time. This incompatibility is addressed by
+margins of safety presented in~\autoref{sec:safetyMargin}.}, the following
+relation holds:
 \begin{equation}
@@
     \item Long--only constraint ensures that a fraction cannot be negative and
-    is added to all candidate companies. Adding this constraint means: "I do not
-    allow short positions.":
+    is added to all candidate companies. It disallows short positions:
     \begin{equation}
@@
     \item Maximum leverage constraint ensures that the leverage is limited up to
-    $L$. Note that $L = 0$ implies no leverage. Adding this constraint means: "I
-    want to limit the use of leverage (to a minimal amount).":
+    $L$. Note that $L = 0$ implies no leverage:
         \begin{equation}
@@
     \item Maximum individual allocation constraint ensures that a fraction does
-    not exceed the specified amount. Adding this constraint means: "I feel
-    uncomfortable putting more than $M$ of my capital into a single company.":
+    not exceed the specified amount. Adding this constraint allows preventing
+    excessive concentration in the portfolio:
         \begin{equation}
@@
     worst--case outcome does not exceed losing a specified amount of capital
-    with a specified probability. Adding this constraint means: "Under a
-    worst--case scenario, I am comfortable permanently losing $K$ (e.g. 25\%) of
-    my capital with probability $P$ (e.g. 0.1\%).":
+    with a specified probability:
         \begin{equation}
@@
     convention, indicating a loss, hence the $\ge$ sign.
+\end{enumerate}
 
@@
 because the fraction $f_j$ is assumed positive. It is also important to note
-that here we do not talk about {\it temporary} loss of capital due to short-term
+that here we do not refer to a {\it temporary} loss of capital due to short--term
 stock market fluctuations, but rather {\it permanent} loss of capital due to the
@@
 course possible for one to lose more than the specified maximum allowable amount
-of capital because of the assumptions made with respect to the inputs: After
-all, coming up with intrinsic values and probabilities for each scenario is more
-of an art than a science.
+of capital because of the assumptions made with respect to the inputs.
 
-\end{enumerate}
-
-\subsection{Constrained System: Putting It All Together}
+\subsection{Constrained System}
 \label{sec:constrainedSystem}
@@
 scenarios (probabilities and intrinsic values) and the same market cap.
-The inputs that present a 50\% loss and 100\% gain with 50--50 probabilities are
-presented in~\autoref{tab:validationCompany}.\\
+The inputs that represent a 50\% loss and 100\% gain with 50--50 probabilities
+are presented in~\autoref{tab:validationCompany}.\\
 
@@
 
-\begin{table}[!ht]
+\begin{table}[!b]
 \caption{Company A with current market cap of 225B USD.}
@@
 
-\begin{table}[!ht]
+\begin{table}[!b]
 \caption{Company B with current market cap of 450M USD.}
@@
 
-\begin{table}[!ht]
+\begin{table}[!b]
 \caption{Company C with current market cap of 39M GBP.}
@@
 
-\begin{table}[!ht]
+\begin{table}[!b]
 \caption{Company D with current market cap of 751M SGD.}
@@
 
-\begin{table}[!ht]
+\begin{table}[!b]
 \caption{Company E with current market cap of 126B HKD.}
@@
 
-\noindent Based on these inputs, with the long--only strategy, without leverage,
+\indent Based on these inputs, with the long--only strategy, without leverage,
 and with maximum individual allocation of 30\%, the portfolio allocation that
@@
 
-\noindent The last item is particularly interesting to the authors. According to
-Actuarial Life Tables in~\cite{lifeTables}, the probability of the (currently 34
-year--old, the oldest) author dying within the next year is approximately
-0.26\%\footnote{This is the only thing that's probably worth measuring in the
-basis points in this framework.}. That is two orders of magnitude higher than
-the probability of the permanent loss of capital for this portfolio. Considering
-that the portfolio has five stocks, that is a very strong argument against
-excessive diversification, especially if:
+\noindent The probability of permanent loss of capital of 0.008\% is
+particularly interesting. Provided that the inputs are reasonably estimated, and
+considering that the portfolio has five stocks, that is a very strong argument
+against excessive diversification, especially if:
 \begin{itemize}
@@
 
-\noindent The observation about excessive diversification is inline with the
+\noindent The observation about excessive diversification is in--line with the
 thoughts from the Poor Charlie's Almanack~\cite{almanack} and one of the
-lectures from Li Lu that the authors frequently watch~\cite{liLuLecture}.
+lectures from Li Lu~\cite{liLuLecture}. In addition, it seems that recent
+research is also drawing similar conclusions about excessive
+diversification~\cite{jamesMenzies,arvanitisEtAl}, although using different
+investing frameworks. The study by James and Menzies~\cite{jamesMenzies} is
+particularly interesting to the authors since the data suggests that strong
+investors should hold concentrated portfolios, which is in--line with practical
+observations of extremely successful fundamentals--based investors: Warren
+Buffett~\cite{berkshireLetters}, Mohnish Pabrai, Nick Sleep~\cite{nomadLetters},
+Pulak Prasad~\cite{darwinInvesting}, and others.
 
@@
 
-\noindent There are a couple of technical problems that the authors observed:
+\noindent A couple of problems have been observed by the authors:
 \begin{enumerate}
@@
     iterations. These errors are ignored, which means that it is possible that
-    the real best solution is not found due to numerical issues.
+    the optimal solution is not found due to numerical issues.
     \item The exponential complexity of the model makes it challenging to
@@
 
-\indent To conclude, the authors believe that the most challenging aspect of an
-investor's work that might use this software is to think hard about the inputs.
-The authors see the usefulness of this software mainly in:
+\indent To conclude, we believe that the most challenging aspect of an
+investor's work that might use this software is to appropriately and
+conservatively model the inputs. The authors see the usefulness of this software
+mainly in:
 \begin{itemize}
@@
     and long--term business outcomes across the range of candidate companies,
-    \item Calculating the optimal allocation in a short amount of time, based on
-    the investor's inputs.
+    \item Avoiding psychological biases by having a tool that calculates the
+    optimal allocation of capital based on a strict mathematical formalism.
 \end{itemize}
@@
 
-\bibliographystyle{unsrt}
-\bibliography{./references}
+\bibliographystyle{biolett}
+\bibliography{./../references}
 

### 2024-02-23 doc/informal/paper.tex
Formal paper (for submission to arXiv)

--- 2024-02-11 doc/paper.tex
@@
 assumptions, the tool calculates the optimal allocation of capital for
-each of the investment candidates. The model has the option to provide
-constraints that ensure: no shorting, no use of leverage, providing a maximum
-limit to the risk of permanent loss of capital, and maximum individual
-allocation. The software is made available for public use.\\
-
-This paper consist of two parts, one without the mathematics and one with the
-mathematics. For a non--mathematical reader, it is best to skip the two sections
-related to mathematics and numerics, namely~\autoref{sec:mathematics}
-and~\autoref{sec:numerics}. The reader who is interested in mathematics should
-still read the preceding sections since they provide context within which the
-mathematics is built on. Moreover, for mathematically skilled reader, the
-problem formulated and solved here represents a non--linear constrained
-optimization problem.
+each of the investment candidates. The model is based on a generalized Kelly
+Criterion with options to provide constraints that ensure: no shorting, no use
+of leverage, providing a maximum limit to the risk of permanent loss of capital,
+and maximum individual allocation. The software is applied to an example
+portfolio from which certain observations about excessive diversification are
+obtained. In addition, the software is made available for public use.\\
 
@@
 The answer to this question has been given by Kelly with his widely known Kelly
-formula (sometimes called Kelly criterion)~\cite{kelly}. The Kelly's approach
-starts by maximizing long--term growth of capital when one is presented with
-an infinite amount of opportunities to bet on. This work extends that idea by
-considering:
+Criterion~\cite{kelly}. Despite the fact that the Kelly Criterion has been
+published in 1956 in the information theory, its use is now widespread
+in the field of finance and sports betting, with a plethora of recent research.
+For example, Pinelis and Ruppret~\cite{pinelisRuppert} noted that the
+asset allocation fractions (weights) of the machine--learning powered Maximally
+Predictable Portfolio (MPP) are similar to weighting by the Kelly criterion.
+Byrnes and Barnett~\cite{byrnesBarnett} went further to generalize the Kelly
+Criterion with a continuous probability distribution modelling the price
+movement of the stock, while Lototsky and Pollok~\cite{lototskyPollok}
+rigorously analyzed the high--frequency limit of the model, concluding that a
+high--frequency strategy can lead to a more aggressive betting strategy.
+Proskurnikov and Barmish~\cite{proskurnikovBarmish} investigated the possibility
+of having a non--linear control for the Kelly betting, assuming a bounded set
+for the input probabilities instead of assuming that they are known in advance.
+The Kelly Criterion was recently even extended to the quantum realm
+by Meister and Price~\cite{meisterPrice}, where in addition to the uncertainty
+related to flipping a coin, there is an uncertainty in measurement. Perhaps the
+most related reference to this work is the paper by Jim\'{e}nez et
+al.~\cite{jimenezEtAl} where they extended the Kelly Criterion to both
+multivariate and simultaneous bets and applied it to sports betting in the
+English Premier League. This work largely follows the approach by Jim\'{e}nez et
+al.~\cite{jimenezEtAl}, but applied to a fundamentals based, focused investing:
 \begin{enumerate}
-    \item Multiple companies (stocks) in parallel,
-    \item An arbitrary number of scenarios for each company,
+    \item Multiple companies (stocks) are considered simultaneously,
+    \item Each with an arbitrary number of scenarios,
     \item Multiple constraints for modelling our preference towards no shorting,
@@
 This work does not try to justify this assumption in a strong mathematical
-form~\footnote{It is on the author's TODO list to try and prove this.}. Here's a
+form~\footnote{It is on the authors' TODO list to try and prove this.}. Here's a
 soft, non-mathematical reasoning on why the authors think this is fine: The
@@
 of an art than a science.
-
-\end{enumerate}
 
@@
 thoughts from the Poor Charlie's Almanack~\cite{almanack} and one of the
-lectures from Li Lu that the authors frequently watch~\cite{liLuLecture}.
+lectures from Li Lu that the authors frequently watch~\cite{liLuLecture}. In
+addition, it seems that recent research is also drawing similar conclusions
+about excessive diversification~\cite{jamesMenzies,arvanitisEtAl}, although
+using different investing frameworks. The study by James and
+Menzies~\cite{jamesMenzies} is particularly interesting to the authors since the
+data suggests that strong investors should hold concentrated portfolios, which
+is in--line with practical observations of famous fundamentals--based investors:
+Warren Buffett~\cite{berkshireLetters}, Mohnish Pabrai, Nick
+Sleep~\cite{nomadLetters}, Pulak Prasad~\cite{darwinInvesting}, and others.
 
